---
title: "Background conversion rate (Bg) across all rounds"
author: "XSun"
date: "2026-08-19"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

# What this is

`Bg_<sample>` is the pipeline's own per-sample estimate of background
(non-m6A) unconverted-read rate -- the mean per-site methylation ratio
(`Uncon/Depth`) across every covered genomic position for that sample, NaNs
dropped (see
[Pipeline overview](pipeline_overview.html#how-bg-background-and-the-binomial-p-value-are-actually-computed)
for the exact formula from `filter_sites.py`). It's the null rate every
site's binomial significance test is measured against.

This page pulls `Bg` for every sample across all 4 rounds into one place, to
check whether the JLK09 elevation we found while investigating round 5
([QC -- round 5](QC_round5_JL_v3.html)) is an isolated case or part of a
broader pattern.

```{r setup, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)

bg_dt <- fread("/project/xinhe/xsun/camseq/2.qc/background_rate_summary.tsv")
bg_dt[, sample := factor(sample, levels = sample[order(round, -Bg)])]
```

# All samples, all rounds

```{r message=FALSE, warning=FALSE, fig.height=5, fig.width=9}
ggplot(bg_dt, aes(x = sample, y = Bg, fill = round)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0.02, linetype = "dashed", color = "grey40") +
  scale_fill_manual(values = c(
    round1_2_PW = "#2a78d6", round3_JL = "#eb6834", round4_JL = "#1baf7a", round5_JL = "#e34948"
  )) +
  labs(
    x = NULL, y = "Bg (background unconverted-read rate)", fill = "Round",
    title = "Background conversion rate by sample",
    caption = "Dashed line at 0.02 for visual reference -- most samples sit just under it."
  ) +
  theme_classic(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r message=FALSE, warning=FALSE}
kable(bg_dt[order(round, -Bg)], digits = c(NA, NA, 5, 0, 0)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

Most samples across all 4 rounds cluster tightly in the 0.013-0.020 range.
Four samples stand out, for three different reasons:

# Case 1: 12-1-3 -- not a real measurement (data failure)

```{r message=FALSE, warning=FALSE}
kable(bg_dt[sample == "12-1-3"]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

`Bg=0` looks clean but isn't -- this sample had only **997 raw reads total**
(vs. tens of millions for every other sample), and only **2 total mapped
positions across the entire 39.4M-position genome table** (max depth at
either position: 1). This is a sequencing/library failure, not a
conversion-rate measurement. Every downstream number for this sample
(including its earlier-reported "0 called sites") is uninterpretable, not
evidence of anything. **Needs its own investigation** -- worth checking
whether the correct FASTQ files were used for this sample, or whether it
needs re-sequencing.

# Case 2 & 3: JL-CAM-01 and 12-1-4 -- input/untreated controls, genuinely elevated

Both have plenty of supporting reads (35.2M and 3.8M deduped genome reads
respectively) -- this is a real signal, not a data artifact. Both are
**input/untreated control samples**, not treated experimental samples.

```{r message=FALSE, warning=FALSE}
kable(bg_dt[sample %in% c("JL-CAM-01", "12-1-4")]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

Checking `Bg` on the rRNA (`genes`) reference too -- a heavily structured
RNA class -- shows the two behave differently, suggesting they may not share
the same root cause despite the similar surface pattern:

```{r message=FALSE, warning=FALSE}
rrna_check <- data.table(
  sample = c("12-1-4", "JL-CAM-01"),
  Bg_genome = c(0.06309, 0.09236),
  Bg_rRNA_genes = c(0.05941, 0.15682)
)
kable(rrna_check, digits = 5) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

- **12-1-4**: similar magnitude in genome vs. rRNA (0.063 vs 0.059) --
  doesn't specifically implicate RNA structure; more consistent with a
  general whole-sample/reaction-level effect.
- **JL-CAM-01**: notably *more* elevated in the structured rRNA reference
  than in the genome (0.157 vs 0.092) -- consistent with secondary structure
  partially blocking chemical conversion access, on top of an already
  elevated baseline.

**Open question for the wet-lab side**: are input/untreated control samples
supposed to skip the chemical A->G conversion treatment step (in which case
elevated background is expected and correct), or should they go through the
same treatment as experimental samples (in which case this indicates a real
technical issue)? That single question resolves whether this pattern is by
design or a problem -- and it's more robust than the JLK09 case below, since
it's consistent across two independent rounds.

# Case 4: JLK09 -- a treated sample, modest but real elevation

```{r message=FALSE, warning=FALSE}
kable(bg_dt[sample %in% c("JLK07", "JLK08", "JLK09")]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

Unlike cases 2-3, JLK09 is a treated experimental sample (Mfs_LPS/RIH007),
not an input/control -- so the "input samples skip conversion" explanation
doesn't apply here. Its elevation is smaller in absolute terms (0.024 vs.
0.06-0.09) but happens in a sample where no obvious protocol difference is
expected, which makes it harder to explain than cases 2-3. See
[QC -- round 5](QC_round5_JL_v3.html) and
[Round 5 miRNA sites](round5_mirna_sites.html) for where this shows up
downstream (JLK09 consistently calls more sites than JLK07/JLK08 across
every cutoff checked, including at miRNA loci specifically).

Debugging directions discussed so far, none of which require new wet-lab
work:

1. **Structured vs. open regions** -- compare `Bg` on `genes` (rRNA) vs
   `genome` for JLK09 specifically, same check as cases 2-3 above (not yet
   run for this sample).
2. **Motif specificity** -- does the excess unconverted signal concentrate
   on real DRACH-like motifs (more consistent with biology) or spread
   uniformly across all A contexts (more consistent with non-specific
   technical noise)? Not yet checked.
3. **Other QC axes** -- base quality, GC content, duplication rate, insert
   size (already in the [MultiQC report](multiqc_round5_JL_v3.html)) for
   JLK09 specifically, to look for any other anomaly pointing to a
   run/lane-specific issue.
4. **A synthetic spike-in control** would resolve this directly and
   definitively, but this experiment doesn't include one -- see the
   discussion in [Pipeline overview](pipeline_overview.html) for what a
   spike-in-based correction would and wouldn't buy us.

# Reference: how Bg is computed

From `filter_sites.py` (see
[Pipeline overview](pipeline_overview.html#how-bg-background-and-the-binomial-p-value-are-actually-computed)
for the full derivation):

```python
df_input.with_columns(
    (pl.col(f"Uncon_{name}") / pl.col(f"Depth_{name}")).alias(f"Ratio_{name}")
    for name in names
).with_columns(
    pl.col(f"Ratio_{name}").drop_nans().mean().alias(f"Bg_{name}")
    for name in names
)
```

An unweighted mean of per-site ratios across every position that sample
covers at all (`Depth > 0`) -- every site counts equally regardless of its
own depth. This is why case 1 (12-1-3) breaks: with only 2 total covered
positions, the "mean" is computed over almost nothing.
