---
title: "Comparing m6a sites from sample5 with camseq paper"
author: "XSun"
date: "2026-02-03"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r message=F, warning=F}

library(data.table)
library(gridExtra)
library(ggplot2)
library(eulerr)
library(GenomicRanges)

# Helper to build file paths
make_file <- function(rd, p_cutoff_str) {
  sprintf(
    "/project/xinhe/xsun/camseq/pipeline/workspace_round3/count_sites/pileup/processed_binomp_%s_1n/sample05_with_id_read%d_binomp%s.tsv",
    p_cutoff_str, rd, p_cutoff_str
  )
}

rd <- 10
pc <- "0.001"
me0 <- 0.05

file_ours <- make_file(rd, pc)
df_ours_raw <- fread(file_ours)

# normalize methylation once; build site once
df_ours_raw[, site := paste0(ref, "-", pos)]

df_ours_f <- df_ours_raw[methylation_mepct >= me0*100]

our_sites <- df_ours_f$site

```


# Introduction

We compare the m6a sites called from sample 2-4 (at p<0.001, methylation level > 0.05, read depth >=10, https://xsun1229.github.io/camseq-m6a/QC_round3.html) with some other databases.

Paper1: [m6A RNA modifications are measured at single-base resolution across the mammalian transcriptome, PMID: 35288668](https://pubmed.ncbi.nlm.nih.gov/35288668/)

The sites are included in Supplementary Data 1, containing the cell types below

| Category | Cell type / Sample | Differentiation stage | RNA preparation | Notes |
|--------|-------------------|----------------------|-----------------|-------|
| Cell line | HeLa | – | poly(A)-selected | Non-immune epithelial cancer cell line |
| Cell line | HEK293 | – | poly(A)-selected | Human embryonic kidney cells |
| Cell line | HEK293 | – | rRNA-depleted (ribo) | Includes pre-mRNA and intronic RNA |
| Cell line | HepG2 | – | poly(A)-selected | Human liver cancer cell line |
| Primary immune | HSPC ribo – d0 | CD34⁺ HSPCs | rRNA-depleted (ribo) | Stem/progenitor stage |
| Primary immune | HSPC ribo – d3 | Early myeloid commitment | rRNA-depleted (ribo) | Early monocytopoiesis |
| Primary immune | HSPC ribo – d6 | Late monocyte progenitors | rRNA-depleted (ribo) | Pre-monocyte stage |
| Primary immune | HSPC ribo – d9 | Monocytes | rRNA-depleted (ribo) | Closest proxy to macrophages |


Paper 2: [REPIC: a database for exploring the N6-methyladenosine methylome, PMID: 32345346](https://pubmed.ncbi.nlm.nih.gov/32345346/)

This paper provides peak-level calls, not single-nucleotide–resolution sites.

data download: https://repicmod.uchicago.edu/repic/download.php

MONO-MAC-6


Paper 3: [m6a-atlas](https://rnamd.org/m6a/merip.php)

peak-level calls, not single-nucleotide–resolution sites.

MONO-MAC-6. 


# Compare with Paper 1, HSPC ribo– d9


```{r message=F, warning=F}

df_paper_raw <- readxl::read_excel("/project/xinhe/xsun/camseq/pipeline/data/m6a_PMID35288668.xlsx", sheet = 8)

df_paper_raw <- as.data.table(df_paper_raw)
df_paper_raw$Chrom <- gsub(df_paper_raw$Chrom, pattern = "chr", replacement = "")
df_paper_raw[, site := paste0(Chrom, "-", Position)]

paper_sites <- df_paper_raw$site

DT::datatable(head(df_paper_raw),caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Several sites reported by camseq paper'),options = list(pageLength = 10) )


print(paste0("Total number of m6a sites reported by the paper ", nrow(df_paper_raw)))

```


```{r message=F, warning=F, fig.height=4, fig.width=5}

rd <- 10
pc <- "0.001"
me0 <- 0.05

inter_sites <- intersect(our_sites, paper_sites)

# merge only the needed columns from paper (smaller/faster)
df_merge <- merge(
  df_ours_f,
  df_paper_raw[, .(site, `m6A fracrion (%)`)],
  by = "site",
  all = FALSE
)

p <-  ggplot(df_merge, aes(x = `m6A fracrion (%)`, y = methylation_mepct)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  #coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    x = "Paper methylation level (%)",
    y = "Our methylation level (%)",
    title = sprintf("read depth %s, p-val %s,\n methylation >= %s", rd, pc, me0)
  ) +
  theme_classic(base_size = 14)
  
fit <- euler(c(
  Paper = length(paper_sites) - length(inter_sites),
  Ours  = length(our_sites) - length(inter_sites),
  "Paper&Ours" = length(inter_sites)
))

p_venn <- plot(
    fit,
    fills = c("#0072B2", "#D55E00"),
    alpha = 0.6,
    labels = TRUE,
    quantities = TRUE,
  )

p

p_venn
```


# Compare with Paper 1, HSPC ribo– d3, HSPC ribo– d6, HSPC ribo– d9


```{r message=F, warning=F}

df_paper_raw1 <- readxl::read_excel("/project/xinhe/xsun/camseq/pipeline/data/m6a_PMID35288668.xlsx", sheet = 6)
df_paper_raw2 <- readxl::read_excel("/project/xinhe/xsun/camseq/pipeline/data/m6a_PMID35288668.xlsx", sheet = 7)
df_paper_raw3 <- readxl::read_excel("/project/xinhe/xsun/camseq/pipeline/data/m6a_PMID35288668.xlsx", sheet = 8)

df_paper_raw <- rbind(df_paper_raw1,df_paper_raw2,df_paper_raw3)

df_paper_raw <- as.data.table(df_paper_raw)
df_paper_raw$Chrom <- gsub(df_paper_raw$Chrom, pattern = "chr", replacement = "")
df_paper_raw[, site := paste0(Chrom, "-", Position)]

df_paper_raw <- df_paper_raw[!duplicated(df_paper_raw$site),]

paper_sites <- df_paper_raw$site

DT::datatable(head(df_paper_raw),caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Several sites reported by camseq paper'),options = list(pageLength = 10) )

print(paste0("Total number of m6a sites reported by the paper ", nrow(df_paper_raw)))

```


```{r message=F, warning=F, fig.height=4, fig.width=5}


inter_sites <- intersect(our_sites, paper_sites)


# merge only the needed columns from paper (smaller/faster)
df_merge <- merge(
  df_ours_f,
  df_paper_raw[, .(site, `m6A fracrion (%)`)],
  by = "site",
  all = FALSE
)

p <-  ggplot(df_merge, aes(x = `m6A fracrion (%)`, y = methylation_mepct)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40") +
  #coord_equal(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    x = "Paper methylation level (%)",
    y = "Our methylation level (%)",
    title = sprintf("read depth %s, p-val %s,\n methylation >= %s", rd, pc, me0)
  ) +
  theme_classic(base_size = 14)
  
fit <- euler(c(
  Paper = length(paper_sites) - length(inter_sites),
  Ours  = length(our_sites) - length(inter_sites),
  "Paper&Ours" = length(inter_sites)
))

p_venn <- plot(
    fit,
    fills = c("#0072B2", "#D55E00"),
    alpha = 0.6,
    labels = TRUE,
    quantities = TRUE,
  )

p

p_venn
```



# Compare with Paper 2, MONO-MAC-6

```{r message=F, warning=F}

df_paper_raw <- fread("/project/xinhe/xsun/camseq/pipeline/data/m6a_REPIC_MONO-MAC-6_hg38.txt")

gr_ours <- GRanges(
  seqnames = paste0("chr", df_ours_f$ref),
  ranges   = IRanges(start = df_ours_f$pos, end = df_ours_f$pos),
  strand   = df_ours_f$strand
)

df_paper <- copy(df_paper_raw)

# remove strand brackets
df_paper[, strand := sub(".*\\[|\\]", "", pos)]

# remove strand part
coord <- sub("\\[.*\\]", "", df_paper$pos)

df_paper[, chr   := sub(":.*", "", coord)]
df_paper[, start := as.integer(sub(".*:(\\d+)-.*", "\\1", coord))]
df_paper[, end   := as.integer(sub(".*-(\\d+)", "\\1", coord))]

gr_paper <- GRanges(
  seqnames = df_paper$chr,
  ranges   = IRanges(start = df_paper$start, end = df_paper$end)
)

hits <- findOverlaps(gr_ours, gr_paper, ignore.strand = FALSE)

n_sites_covered <- length(unique(queryHits(hits)))
n_sites_total   <- length(gr_ours)

df_bar <- data.table(
  status = c("Covered by REPIC peak", "Not covered"),
  count  = c(
    n_sites_covered,
    n_sites_total - n_sites_covered
  )
)

df_bar[, percent := count / n_sites_total * 100]

p_bar <- ggplot(df_bar, aes(x = status, y = count, fill = status)) +
  geom_col(width = 0.6, alpha = 0.8) +
  geom_text(
    aes(label = sprintf("%d\n(%.1f%%)", count, percent)),
    vjust = -0.4,
    size = 4
  ) +
  scale_fill_manual(values = c("#0072B2", "grey70")) +
  labs(
    y = "Number of sites",
    x = NULL,
    title = "Coverage of our sites by REPIC m6A peaks"
  ) +
  theme_classic(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5)
  )

p_bar

```



# Compare with Paper 3, MONO-MAC-6

```{r message=F, warning=F}

df_paper_raw <- fread("/project/xinhe/xsun/camseq/pipeline/data/m6a_Atlas2_MONO-MAC-6.txt")

gr_ours <- GRanges(
  seqnames = paste0("chr", df_ours_f$ref),
  ranges   = IRanges(start = df_ours_f$pos, end = df_ours_f$pos),
  strand   = df_ours_f$strand
)

gr_paper <- GRanges(
  seqnames = df_paper_raw$seqnames,
  ranges   = IRanges(df_paper_raw$start, df_paper_raw$end),
  strand   = df_paper_raw$strand
)

hits <- findOverlaps(gr_ours, gr_paper, ignore.strand = FALSE)
n_sites_covered <- length(unique(queryHits(hits)))
n_sites_total   <- length(gr_ours)
n_sites_covered / n_sites_total

df_bar <- data.table(
  status = c("Covered by m6A-Atlas peak", "Not covered"),
  count  = c(n_sites_covered, n_sites_total - n_sites_covered)
)
df_bar[, percent := 100 * count / n_sites_total]

p_bar <- ggplot(df_bar, aes(x = status, y = count, fill = status)) +
  geom_col(width = 0.6, alpha = 0.85) +
  geom_text(
    aes(label = sprintf("%s\n(%.1f%%)", format(count, big.mark=","), percent)),
    vjust = -0.4,
    size = 4
  ) +
  scale_fill_manual(values = c("#0072B2", "grey70")) +
  labs(x = NULL, y = "Number of sites",
       title = "Coverage of our sites by m6A-Atlas peak") +
  theme_classic(base_size = 14) +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

p_bar

```

