---
title: "Summary after processing"
author: "XSun"
date: "2025-03-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
library(ggplot2)
library(data.table)
library(tidyr)
```

# ```{r fig.height=8, fig.width=16}
# 
# stat_folder <- "/project/xinhe/xsun/camseq/data/workspace/internal_files/aligned_bam/stat/"
# stat_files_gc_detail <- list.files(stat_folder, pattern = "*gc_metrics", full.names = TRUE)
# stat_files_gc_summary <- list.files(stat_folder, pattern = "*gc_summary", full.names = TRUE)
# 
# p <- list()
# for (i in 1:length(stat_files_gc_detail)){
#   
#   file_gc_detail <- stat_files_gc_detail[i]
#   file_gc_summary <- stat_files_gc_summary[i]
#   
#   sample_name <- tools::file_path_sans_ext(basename(file_gc_detail))  # Get sample name from file name
#   sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
#   sample_name <- gsub(pattern = ".gc_metrics",replacement = "",x =sample_name)
#   
#   detailed <- read.delim(file_gc_detail, 
#                      comment.char = "#", 
#                      stringsAsFactors = FALSE)
#   
#   p[[length(p)+1]] <- ggplot(detailed, aes(x = GC, y = NORMALIZED_COVERAGE)) +
#   geom_line(color = "steelblue", linewidth = 1) +
#   geom_point(color = "navy", size = 1.5) +
#   labs(title = sample_name,
#        x = "GC Percentage",
#        y = "Normalized Coverage") +
#   theme_minimal() +
#   theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
#   scale_x_continuous(breaks = seq(0, 100, 10))
#   
#   
#   summary <- read.delim(file_gc_summary, 
#                      comment.char = "#", 
#                      stringsAsFactors = FALSE)
#   
#   summary_long <- summary %>%
#   select(starts_with("GC_NC")) %>%
#   pivot_longer(cols = everything(), 
#               names_to = "GC_Bin", 
#               values_to = "Normalized_Coverage") %>%
#   mutate(GC_Bin = gsub("GC_NC_", "", GC_Bin),
#          GC_Bin = gsub("_", "-", GC_Bin))
#   
# }
# 
# combined <- arrangeGrob(
#   grobs = p,
#   nrow = 2,
#   top = textGrob("GC Bias Profile", gp = gpar(fontsize = 14))
# )
# grid.draw(combined)
# 
# ```



## Alignment Statistics

```{r fig.height=8, fig.width=16}

stat_folder <- "/project/xinhe/xsun/camseq/data/workspace/internal_files/aligned_bam/stat/"
stat_files <- list.files(stat_folder, pattern = "*flagstat", full.names = TRUE)

p <- list()
for (file in stat_files){
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  sample_name <- gsub(pattern = ".genome",replacement = "",x =sample_name)
  
  flagstat_lines <- readLines(file) 
  flagstat_data <- data.frame(
  Metric = c("Total Reads", "Mapped Reads", "Properly Paired", "Duplicates"),
  Value = c(
    as.numeric(gsub(" .*", "", flagstat_lines[1])),        # Total reads
    as.numeric(gsub(" .*", "", flagstat_lines[7])),        # Mapped reads
    as.numeric(gsub(" .*", "", flagstat_lines[11])),       # Properly paired
    as.numeric(gsub(" .*", "", flagstat_lines[4]))         # Duplicates
    )
  )
  
  flagstat_data$Metric <- factor(flagstat_data$Metric,
                              levels = c("Total Reads", "Mapped Reads", 
                                       "Properly Paired", "Duplicates"))

  p[[length(p)+1]] <- ggplot(flagstat_data, aes(x = Metric, y = Value, fill = Metric)) +
  geom_bar(stat = "identity") +
  labs(title = sample_name, y = "Count") +
  theme_minimal()
  
}

combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("Alignment Statistics", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```


## Insert Size Distribution

```{r fig.height=8, fig.width=16}

stat_folder <- "/project/xinhe/xsun/camseq/data/workspace/internal_files/aligned_bam/stat/"
stat_files <- list.files(stat_folder, pattern = "*insert_metrics.txt", full.names = TRUE)

p <- list()
for (file in stat_files){
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  sample_name <- gsub(pattern = ".insert_metrics",replacement = "",x =sample_name)
  
  insert_size <- read.table(file, 
                          skip = 10, header = TRUE, comment.char = "#", nrows = 1000)
  
  p[[length(p)+1]] <- ggplot(insert_size, aes(x = insert_size, y = All_Reads.fr_count)) +
                      geom_bar(stat = "identity", fill = "purple") +
                      labs(x = "Insert Size (bp)", y = "Frequency", title = sample_name)
  
}

combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("Insert Size Distribution", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

## Reads per Chromosome

```{r fig.height=8, fig.width=32}

stat_folder <- "/project/xinhe/xsun/camseq/data/workspace/internal_files/aligned_bam/stat/"
stat_files <- list.files(stat_folder, pattern = "*chrom_counts.txt", full.names = TRUE)

p <- list()
for (file in stat_files){
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  sample_name <- gsub(pattern = ".genome.chrom_counts",replacement = "",x =sample_name)
  
  chrom_counts <- read.table(file, 
                          col.names = c("chr", "length", "mapped", "unmapped"))
  
  chrom_counts <- chrom_counts[!grepl("_", chrom_counts$chr), ]
  # chrom_counts <- chrom_counts[chrom_counts$chr %in% c(1:22),]
  # chrom_counts <- chrom_counts[order(as.numeric(chrom_counts$chr), decreasing = F),]
  # chrom_counts$chr <- factor(chrom_counts$chr,levels = c(1:22))
  
  # p[[length(p)+1]] <- ggplot(chrom_counts, aes(x = chr, y = mapped)) +
  #                     geom_bar(stat = "identity") +
  #                     labs(x = "Chromosome", y = "Read Counts", title = sample_name) +
  #                     theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
  p[[length(p)+1]] <- ggplot(chrom_counts, aes(x = reorder(chr,-mapped), y = mapped)) +
                      geom_bar(stat = "identity") +
                      labs(x = "Chromosome", y = "Read Counts", title = sample_name) +
                      theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
}

combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("Reads per Chromosome", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

## MAPQ Distribution

```{r fig.height=8, fig.width=16}

stat_folder <- "/project/xinhe/xsun/camseq/data/workspace/internal_files/aligned_bam/stat/"
stat_files <- list.files(stat_folder, pattern = "*mapq_dist.txt", full.names = TRUE)

p <- list()
for (file in stat_files){
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  sample_name <- gsub(pattern = "genome.mapq_dist",replacement = "",x =sample_name)
  
  mapq <- read.table(file, col.names = c("Count", "MAPQ"))
  
  mapq$MAPQ <- factor(mapq$MAPQ, levels = c(1, 60), labels = c("1", "60"))
  
  p[[length(p)+1]] <- ggplot(mapq, aes(x = MAPQ, y = Count)) +
  geom_bar(stat = "identity", fill = "darkgreen", width = 0.6) +  # Adjust width
  labs(
    title = sample_name,
    x = "Mapping Quality",
    y = "Count"
  ) +
  theme_minimal()
  
}

combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("MAPQ Distribution", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```


## Read Length Distribution

```{r fig.height=8, fig.width=16}

stat_folder <- "/project/xinhe/xsun/camseq/data/workspace/internal_files/aligned_bam/stat/"
stat_files <- list.files(stat_folder, pattern = "*read_lengths.txt", full.names = TRUE)

p <- list()
for (file in stat_files){
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  sample_name <- gsub(pattern = "genome.read_lengths",replacement = "",x =sample_name)
  
 
  read_lengths <- read.table(file, col.names = c("Count", "Length"))


  p[[length(p)+1]] <- ggplot(read_lengths, aes(x = Length, y = Count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = sample_name, x = "Read Length (bp)", y = "Count")
  
}

combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("Read Length Distribution", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```



gc_bias <- read.table("sample.gc_bias.txt", skip = 6, header = TRUE, comment.char = "#")

ggplot(gc_bias, aes(x = GC, y = NORMALIZED_COVERAGE)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "GC Bias", x = "GC%", y = "Normalized Coverage") +
  geom_hline(yintercept = 1, linetype = "dashed")