---
title: "Quality Control"
author: "XSun"
date: "2025-03-17"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r message=F, warning=F}

library(jsonlite)
library(dplyr)
library(gridExtra)
library(ggplot2)
library(grid) 
library(stringr)
library(tidyverse)
library(ggsci)

# Function to extract numeric values using regex
extract_value <- function(pattern) {
  line <- str_subset(lines, pattern)
  as.numeric(str_extract(line, "\\d+"))
}
```

# Sample info 

| Sample Name |         | i7 Index Sequence   |         | i5 Index Sequence   | i5 reverse complement | Index Cat no | Species            | Library type (DNA or RNA, mutation or not) | ng/ul | data      |
|-------------|---------|---------------------|---------|---------------------|-----------------------|--------------|--------------------|--------------------------------------------|-------|-----------|
| 12-1-1      | 701     | ATTACTCG            | 501     | TATAGCCT            | AGGCTATA              | NEB#E7600    | 18-QTL treatment   | CAM-seq (A to G)                           | 0.246 | 490,332   |
| 12-1-2      | 702     | TCCGGAGA            | 501     | TATAGCCT            | AGGCTATA              | NEB#E7600    | 28-QTL treatment   | CAM-seq (A to G)                           | 0.376 | 2,367,054 |
| 12-1-3      | 713     | TTCCTCCT            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 18-QTL untreatment | CAM-seq (A to G)                           | 0.824 | 1,590,686 |
| 12-1-4      | 714     | TGCTTGCT            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 28-QTL untreatment | CAM-seq (A to G)                           | 1.03  | 1,898,213 |
| 12-1-5      | 715     | GGTGATGA            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 18-QTL untreatment | CAM-seq (A to G)                           | 0.268 | 272,739   |
| 12-1-6      | 716     | AACCTACG            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 28-QTL untreatment | CAM-seq (A to G)                           | <0.1  | 659,559   |
| 12-1-7      | 721&724 | AAGGCGTA & AATCCAGC | 515&516 | TGTGGCAT & ACTACGGA | ATGCCACA & TCCGTAGT   | NEB#E7780    | 18-QTL treatment   | CAM-seq (A to G)                           | 0.25  | 870,124   |
| 12-1-8      | 722     | CTGATGAG            | 515&516 | TGTGGCAT & ACTACGGA | ATGCCACA & TCCGTAGT   | NEB#E7780    | 28-QTL treatment   | CAM-seq (A to G)                           | 0.29  | 960,388   |

# Output from multiqc 

## Pipeline output 

https://xsun1229.github.io/camseq-m6a/multiqc_report_genome_all.html

## QC for the aligned bam files

https://xsun1229.github.io/camseq-m6a/multiqc_report_aligned_bam.html


# Trimming

```{r}
json_folder <- "/project/xinhe/xsun/camseq/data/workspace/report_reads/trimming/"

# List all JSON files in the folder
json_files <- list.files(json_folder, pattern = "*.json", full.names = TRUE)

avg_length_all <- c()
for (file in json_files) {
  # Read the JSON file
  
  json_data <- fromJSON(file)
  
  total_bases <- json_data$basepair_counts$output
  output_reads <- json_data$read_counts$output
  
  avg_length <- round(total_bases / (output_reads * 2),digits = 4)  # Paired-end
  sample <- extracted_pattern <- sub(".*24JX-([0-9-]+)_run1.*", "\\1", file)
  
  tmp <- c(sample, avg_length)
  avg_length_all <- rbind(avg_length_all,tmp)
  
}

rownames(avg_length_all) <- NULL
colnames(avg_length_all) <- c("sample","avg_length/read")
avg_length_all <- as.data.frame(avg_length_all )

DT::datatable(avg_length_all,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Average length after trimming'),options = list(pageLength = 10) )
```


# Mythelytion level

## cutoff: covarage = 10, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=5}

counts <- 10
pct <- 10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/"

file_list <- list.files(path = folder_countsites,pattern = paste0("counts_cutoff",counts,"_pct",pct),full.names = T)
file_list <- file_list[c(1,2,7,8)]
combined_data <- map_df(file_list, ~{
  # Extract sample name from filename
  sample_name <- str_remove(.x, "_motif_counts_cutoff\\d+_pct\\d+.tsv$") %>%
    str_remove(".*241217AVITI-He-lab-24JX-") %>%  # Remove prefix
    str_remove("^/+") %>%  # Remove leading slashes if present
    str_remove("/+$")       # Remove trailing slashes if present
  
  read_tsv(.x) %>%
    filter(motif != "motif") %>%  # Remove header remnants
    mutate(
      sample = sample_name,
      percentage = (count / sum(count)) * 100
    ) %>%
    select(sample, motif, percentage)
})

combined_data <- combined_data %>%
  mutate(motif = str_replace(motif, "(.)([+-])(.)", "\\2\\1A\\3"))

combined_data_gac <- combined_data[combined_data$motif %in% c("+GAC","-GAC"),]
combined_data_gac<-  combined_data_gac %>%
  group_by(sample) %>%
  mutate(total_per_sample = sum(percentage)) %>%
  ungroup()  # Optional: removes grouping
combined_data_aac <- combined_data[combined_data$motif %in% c("+AAC","-AAC"),]
combined_data_aac <- combined_data_aac %>%
  group_by(sample) %>%
  mutate(total_per_sample = sum(percentage)) %>%
  ungroup()  # Optional: removes grouping

combined_data_gac$total_per_sample[seq(2,nrow(combined_data_gac),by=2)] <- "-"
combined_data_aac$total_per_sample[seq(2,nrow(combined_data_aac),by=2)] <- "-"
aac_gac <- cbind(combined_data_gac,combined_data_aac)

DT::datatable(aac_gac,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Percentage for AAC,GAC motif'),options = list(pageLength = 10) )


# Create visualization
ggplot(combined_data, aes(x = percentage, y = sample)) +
  geom_point(
    aes(color = motif),
    size = 3,
    position = position_jitterdodge(
      jitter.width = 0.2,
      jitter.height = 0.1,
      dodge.width = 0.7
    ),
    show.legend = FALSE  # <--- This removes the point legend
  ) +
  geom_text(
    aes(label = motif, color = motif),
    position = position_jitterdodge(
      jitter.width = 0.2,
      jitter.height = 0.1,
      dodge.width = 0.7
    ),
    size = 3,
    vjust = -1.5,
    show.legend = FALSE  # <--- This removes the text legend
  ) +
  scale_color_igv(guide = "none") +  # <--- Disables color scale legend
  labs(
    title = "Motif Percentage Distribution Across Samples",
    x = "Percentage (%)",
    y = "Sample",
    caption = paste("-/+ represent strands")
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",  # Ensures no residual legend elements
    panel.grid.major.y = element_line(color = "grey90"),
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```


## cutoff: covarage = 10, methelytion level = 20%

```{r warning=F, message=F, fig.height=5, fig.width=5}

counts <- 10
pct <- 20

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/"

file_list <- list.files(path = folder_countsites,pattern = paste0("counts_cutoff",counts,"_pct",pct),full.names = T)
file_list <- file_list[c(1,2,7,8)]
combined_data <- map_df(file_list, ~{
  # Extract sample name from filename
  sample_name <- str_remove(.x, "_motif_counts_cutoff\\d+_pct\\d+.tsv$") %>%
    str_remove(".*241217AVITI-He-lab-24JX-") %>%  # Remove prefix
    str_remove("^/+") %>%  # Remove leading slashes if present
    str_remove("/+$")       # Remove trailing slashes if present
  
  read_tsv(.x) %>%
    filter(motif != "motif") %>%  # Remove header remnants
    mutate(
      sample = sample_name,
      percentage = (count / sum(count)) * 100
    ) %>%
    select(sample, motif, percentage)
})

combined_data <- combined_data %>%
  mutate(motif = str_replace(motif, "(.)([+-])(.)", "\\2\\1A\\3"))

combined_data_gac <- combined_data[combined_data$motif %in% c("+GAC","-GAC"),]
combined_data_gac<-  combined_data_gac %>%
  group_by(sample) %>%
  mutate(total_per_sample = sum(percentage)) %>%
  ungroup()  # Optional: removes grouping
combined_data_aac <- combined_data[combined_data$motif %in% c("+AAC","-AAC"),]
combined_data_aac <- combined_data_aac %>%
  group_by(sample) %>%
  mutate(total_per_sample = sum(percentage)) %>%
  ungroup()  # Optional: removes grouping

combined_data_gac$total_per_sample[seq(2,nrow(combined_data_gac),by=2)] <- "-"
combined_data_aac$total_per_sample[seq(2,nrow(combined_data_aac),by=2)] <- "-"
aac_gac <- cbind(combined_data_gac,combined_data_aac)

DT::datatable(aac_gac,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Percentage for AAC,GAC motif'),options = list(pageLength = 10) )


# Create visualization
ggplot(combined_data, aes(x = percentage, y = sample)) +
  geom_point(
    aes(color = motif),
    size = 3,
    position = position_jitterdodge(
      jitter.width = 0.2,
      jitter.height = 0.1,
      dodge.width = 0.7
    ),
    show.legend = FALSE  # <--- This removes the point legend
  ) +
  geom_text(
    aes(label = motif, color = motif),
    position = position_jitterdodge(
      jitter.width = 0.2,
      jitter.height = 0.1,
      dodge.width = 0.7
    ),
    size = 3,
    vjust = -1.5,
    show.legend = FALSE  # <--- This removes the text legend
  ) +
  scale_color_igv(guide = "none") +  # <--- Disables color scale legend
  labs(
    title = "Motif Percentage Distribution Across Samples",
    x = "Percentage (%)",
    y = "Sample",
    caption = paste("-/+ represent strands")
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",  # Ensures no residual legend elements
    panel.grid.major.y = element_line(color = "grey90"),
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```


## cutoff: covarage = 10, methelytion level = 30%


```{r warning=F, message=F, fig.height=5, fig.width=5}

counts <- 10
pct <- 30

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/"

file_list <- list.files(path = folder_countsites,pattern = paste0("counts_cutoff",counts,"_pct",pct),full.names = T)
file_list <- file_list[c(1,2,7,8)]
combined_data <- map_df(file_list, ~{
  # Extract sample name from filename
  sample_name <- str_remove(.x, "_motif_counts_cutoff\\d+_pct\\d+.tsv$") %>%
    str_remove(".*241217AVITI-He-lab-24JX-") %>%  # Remove prefix
    str_remove("^/+") %>%  # Remove leading slashes if present
    str_remove("/+$")       # Remove trailing slashes if present
  
  read_tsv(.x) %>%
    filter(motif != "motif") %>%  # Remove header remnants
    mutate(
      sample = sample_name,
      percentage = (count / sum(count)) * 100
    ) %>%
    select(sample, motif, percentage)
})

combined_data <- combined_data %>%
  mutate(motif = str_replace(motif, "(.)([+-])(.)", "\\2\\1A\\3"))

combined_data_gac <- combined_data[combined_data$motif %in% c("+GAC","-GAC"),]
combined_data_gac<-  combined_data_gac %>%
  group_by(sample) %>%
  mutate(total_per_sample = sum(percentage)) %>%
  ungroup()  # Optional: removes grouping
combined_data_aac <- combined_data[combined_data$motif %in% c("+AAC","-AAC"),]
combined_data_aac <- combined_data_aac %>%
  group_by(sample) %>%
  mutate(total_per_sample = sum(percentage)) %>%
  ungroup()  # Optional: removes grouping

combined_data_gac$total_per_sample[seq(2,nrow(combined_data_gac),by=2)] <- "-"
combined_data_aac$total_per_sample[seq(2,nrow(combined_data_aac),by=2)] <- "-"
aac_gac <- cbind(combined_data_gac,combined_data_aac)

DT::datatable(aac_gac,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Percentage for AAC,GAC motif'),options = list(pageLength = 10) )


# Create visualization
ggplot(combined_data, aes(x = percentage, y = sample)) +
  geom_point(
    aes(color = motif),
    size = 3,
    position = position_jitterdodge(
      jitter.width = 0.2,
      jitter.height = 0.1,
      dodge.width = 0.7
    ),
    show.legend = FALSE  # <--- This removes the point legend
  ) +
  geom_text(
    aes(label = motif, color = motif),
    position = position_jitterdodge(
      jitter.width = 0.2,
      jitter.height = 0.1,
      dodge.width = 0.7
    ),
    size = 3,
    vjust = -1.5,
    show.legend = FALSE  # <--- This removes the text legend
  ) +
  scale_color_igv(guide = "none") +  # <--- Disables color scale legend
  labs(
    title = "Motif Percentage Distribution Across Samples",
    x = "Percentage (%)",
    y = "Sample",
    caption = paste("-/+ represent strands")
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",  # Ensures no residual legend elements
    panel.grid.major.y = element_line(color = "grey90"),
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```