---
title: "Quality Control"
author: "XSun"
date: "2025-03-17"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r message=F, warning=F}

library(jsonlite)
library(dplyr)
library(gridExtra)
library(ggplot2)
library(grid) 
library(stringr)


# Function to extract numeric values using regex
extract_value <- function(pattern) {
  line <- str_subset(lines, pattern)
  as.numeric(str_extract(line, "\\d+"))
}
```

# Sample info 

| Sample Name |         | i7 Index Sequence   |         | i5 Index Sequence   | i5 reverse complement | Index Cat no | Species            | Library type (DNA or RNA, mutation or not) | ng/ul | data      |
|-------------|---------|---------------------|---------|---------------------|-----------------------|--------------|--------------------|--------------------------------------------|-------|-----------|
| 12-1-1      | 701     | ATTACTCG            | 501     | TATAGCCT            | AGGCTATA              | NEB#E7600    | 18-QTL treatment   | CAM-seq (A to G)                           | 0.246 | 490,332   |
| 12-1-2      | 702     | TCCGGAGA            | 501     | TATAGCCT            | AGGCTATA              | NEB#E7600    | 28-QTL treatment   | CAM-seq (A to G)                           | 0.376 | 2,367,054 |
| 12-1-3      | 713     | TTCCTCCT            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 18-QTL untreatment | CAM-seq (A to G)                           | 0.824 | 1,590,686 |
| 12-1-4      | 714     | TGCTTGCT            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 28-QTL untreatment | CAM-seq (A to G)                           | 1.03  | 1,898,213 |
| 12-1-5      | 715     | GGTGATGA            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 18-QTL untreatment | CAM-seq (A to G)                           | 0.268 | 272,739   |
| 12-1-6      | 716     | AACCTACG            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 28-QTL untreatment | CAM-seq (A to G)                           | <0.1  | 659,559   |
| 12-1-7      | 721&724 | AAGGCGTA & AATCCAGC | 515&516 | TGTGGCAT & ACTACGGA | ATGCCACA & TCCGTAGT   | NEB#E7780    | 18-QTL treatment   | CAM-seq (A to G)                           | 0.25  | 870,124   |
| 12-1-8      | 722     | CTGATGAG            | 515&516 | TGTGGCAT & ACTACGGA | ATGCCACA & TCCGTAGT   | NEB#E7780    | 28-QTL treatment   | CAM-seq (A to G)                           | 0.29  | 960,388   |

# Trimming

https://xsun1229.github.io/camseq-m6a/multiqc_report_trimming.html


```{r}
json_folder <- "/project/xinhe/xsun/camseq/data/workspace/report_reads/trimming/"

# List all JSON files in the folder
json_files <- list.files(json_folder, pattern = "*.json", full.names = TRUE)

avg_length_all <- c()
for (file in json_files) {
  # Read the JSON file
  
  json_data <- fromJSON(file)
  
  total_bases <- json_data$basepair_counts$output
  output_reads <- json_data$read_counts$output
  
  avg_length <- round(total_bases / (output_reads * 2),digits = 4)  # Paired-end
  sample <- extracted_pattern <- sub(".*24JX-([0-9-]+)_run1.*", "\\1", file)
  
  tmp <- c(sample, avg_length)
  avg_length_all <- rbind(avg_length_all,tmp)
  
}

rownames(avg_length_all) <- NULL
colnames(avg_length_all) <- c("sample","avg_length/read")
avg_length_all <- as.data.frame(avg_length_all )

DT::datatable(avg_length_all,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Average length after trimming'),options = list(pageLength = 10) )
```

# Mapping to genome


https://xsun1229.github.io/camseq-m6a/multiqc_report_mapping.html


# Deduplicating 


https://xsun1229.github.io/camseq-m6a/multiqc_report_dedup.html


# Aligned bam QC

https://xsun1229.github.io/camseq-m6a/multiqc_report_aligned_bam.html

# Mythelytion level

Remove low-coverage sites

```{r warning=F, message=F, fig.height=8, fig.width=16}

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/"
files_countsites <- list.files(path = folder_countsites,pattern = "genome.tsv",full.names = T)

p_me_dist <- list()
p_cov_dist <- list()
strand_bias_all <- c()
for (file in files_countsites) {
  
  sample <- regmatches(basename(file),
            regexpr("\\d+-\\d+-\\d+(?=\\.)", basename(file), perl = TRUE))
  
  df <- data.table::fread(file)
  
  # Calculate methylation and filter
  df <- df %>%
    mutate(
      total = convertedBaseCount + unconvertedBaseCount,
      methylation_pct = (unconvertedBaseCount / total) * 100
    ) %>%
    filter(total >= 10)   # Remove low-coverage sites
  
  if (nrow(df) > 0) {
    strand_bias <- df %>%
    group_by(strand) %>%
    summarise(
      mean_methylation = round(mean(methylation_pct, na.rm = TRUE),digits = 4),
      n_sites_pass_filtering = n()
    ) %>%
    mutate(sample = sample) %>%          # Add sample column
    select(sample, everything())         # Reorder columns (sample first)
  }else{
     strand_bias <- data.frame(sample = c(sample,sample),
                               strand = c("+","-"),
                               mean_methylation = c("no site passing filtering","no site passing filtering"),
                               n_sites_pass_filtering = c("no site passing filtering","no site passing filtering"))
    
  }
  
  strand_bias_all <- rbind(strand_bias_all,strand_bias)
  
  p_me_dist[[sample]] <- ggplot(df, aes(x=methylation_pct)) +
    geom_histogram(bins=30, fill="blue", alpha = 0.6) +
    labs(x="Methylation (%)", y="Number of sites") +
    ggtitle(sample) +
    theme_minimal()
  
  p_cov_dist[[sample]] <- ggplot(df, aes(x=total)) +
    geom_histogram(bins=50, fill="blue", alpha = 0.6) +
    #xlim(-1, 100) +
    labs(x="Total reads per site", y="Count") +
    ggtitle(sample) +
    theme_minimal()


  
  
  
}


combined <- arrangeGrob(
  grobs = p_me_dist,
  nrow = 2,
  top = textGrob("Methylation Distribution", gp = gpar(fontsize = 14))
)
grid.draw(combined)
```

```{r warning=F, message=F, fig.height=8, fig.width=16}
combined <- arrangeGrob(
  grobs = p_cov_dist,
  nrow = 2,
  top = textGrob("Coverage Distribution", gp = gpar(fontsize = 14))
)
grid.draw(combined)


DT::datatable(strand_bias_all,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Checking the strand bias '),options = list(pageLength = 10) )

```



