---
title: "Quality Control"
author: "XSun"
date: "2025-03-17"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r message=F, warning=F}

library(jsonlite)
library(dplyr)
library(gridExtra)
library(ggplot2)
library(grid) 
library(stringr)
library(tidyverse)
library(ggsci)

library(tidyverse)
library(readr)
library(purrr)
library(stringr)
library(ggplot2)

process_motif_data <- function(file_list) {
  # Process files and calculate totals
  combined_data <- map_df(file_list, ~{
    message("\nProcessing: ", normalizePath(.x))
    
    # Extract sample name
    sample_name <- basename(.x) %>%
      str_remove("_motif_summary_cov\\d+_mepct\\d+\\.tsv$") %>%
      str_remove("241217AVITI-He-lab-24JX-") %>%
      str_remove("^_+") %>%
      str_remove("_+$")
    
    # Read and process data
    sample_data <- read_table(.x, 
                             col_names = c("motif", "count"), 
                             col_types = "cn",
                             skip_empty_rows = TRUE)
    
    total_reads <- sum(sample_data$count)
    
    processed <- sample_data %>%
      mutate(
        sample = sample_name,
        total_count = total_reads,
        percentage = (count / total_reads) * 100
      ) %>%
      select(sample, motif, percentage, total_count)
    
    return(processed)
  })
  
  # Create formatted labels and order data
  combined_data %>%
    mutate(
      sample_label = paste0(sample, " (n=", 
                           format(total_count, big.mark = ",", scientific = FALSE), ")"),
      sample_label = fct_reorder(sample_label, total_count)
    ) %>%
    arrange(desc(total_count)) %>%
    mutate(sample_label = factor(sample_label, levels = unique(sample_label)))
}

create_motif_plot <- function(processed_motif_data, 
                             title = "Motif Distribution with Total Read coverage",
                             x_lab = "Percentage (%)",
                             y_lab = "Sample (Total site coverage passing filter)",
                             point_size = 3,
                             text_size = 3) {
  ggplot(processed_motif_data, aes(x = percentage, y = sample_label)) +
    geom_point(
      aes(color = motif),
      size = point_size,
      position = position_jitterdodge(
        jitter.width = 0.2,
        jitter.height = 0.1,
        dodge.width = 0.7
      ),
      show.legend = FALSE
    ) +
    geom_text(
      aes(label = motif, color = motif),
      position = position_jitterdodge(
        jitter.width = 0.2,
        jitter.height = 0.1,
        dodge.width = 0.7
      ),
      size = text_size,
      vjust = -1.5,
      show.legend = FALSE
    ) +
    scale_color_igv(guide = "none") +
    labs(
      title = title,
      x = x_lab,
      y = y_lab
    ) +
    theme_minimal() +
    theme(
      panel.grid.major.y = element_line(color = "grey90"),
      axis.text.y = element_text(size = 9, face = "bold"),
      axis.text.x = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 14)
    )
}

count_overlap <- function(file_list) {
  # Initialize an empty list to store the 'name' columns
  name_list <- list()
  
  # Read each file and store the 'name' column
  for(file in file_list){
    df_site <- read.table(file, header = T)
    df_site$name <- paste0(df_site$ref, "_", df_site$pos)
    name_list <- append(name_list, list(df_site$name))
  }
  
  # Find the overlap among all 'name' columns
  overlap <- Reduce(intersect, name_list)
  
  # Return the overlap count
  return(length(overlap))
}


```

# Sample info 

| Sample Name |         | i7 Index Sequence   |         | i5 Index Sequence   | i5 reverse complement | Index Cat no | Species            | Library type (DNA or RNA, mutation or not) | ng/ul | data      |
|-------------|---------|---------------------|---------|---------------------|-----------------------|--------------|--------------------|--------------------------------------------|-------|-----------|
| 12-1-1      | 701     | ATTACTCG            | 501     | TATAGCCT            | AGGCTATA              | NEB#E7600    | 18-QTL treatment   | CAM-seq (A to G)                           | 0.246 | 490,332   |
| 12-1-2      | 702     | TCCGGAGA            | 501     | TATAGCCT            | AGGCTATA              | NEB#E7600    | 28-QTL treatment   | CAM-seq (A to G)                           | 0.376 | 2,367,054 |
| 12-1-3      | 713     | TTCCTCCT            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 18-QTL untreatment | CAM-seq (A to G)                           | 0.824 | 1,590,686 |
| 12-1-4      | 714     | TGCTTGCT            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 28-QTL untreatment | CAM-seq (A to G)                           | 1.03  | 1,898,213 |
| 12-1-5      | 715     | GGTGATGA            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 18-QTL untreatment | CAM-seq (A to G)                           | 0.268 | 272,739   |
| 12-1-6      | 716     | AACCTACG            | 509     | TTGCTTGC            | GCAAGCAA              | NEB#E7780    | 28-QTL untreatment | CAM-seq (A to G)                           | <0.1  | 659,559   |
| 12-1-7      | 721&724 | AAGGCGTA & AATCCAGC | 515&516 | TGTGGCAT & ACTACGGA | ATGCCACA & TCCGTAGT   | NEB#E7780    | 18-QTL treatment   | CAM-seq (A to G)                           | 0.25  | 870,124   |
| 12-1-8      | 722     | CTGATGAG            | 515&516 | TGTGGCAT & ACTACGGA | ATGCCACA & TCCGTAGT   | NEB#E7780    | 28-QTL treatment   | CAM-seq (A to G)                           | 0.29  | 960,388   |

# Output from multiqc 

## Pipeline output 

https://xsun1229.github.io/camseq-m6a/multiqc_report_genome_all.html

## QC for the aligned bam files

https://xsun1229.github.io/camseq-m6a/multiqc_report_aligned_bam.html


# Trimming

```{r}
json_folder <- "/project/xinhe/xsun/camseq/data/workspace/report_reads/trimming/"

# List all JSON files in the folder
json_files <- list.files(json_folder, pattern = "*.json", full.names = TRUE)

avg_length_all <- c()
for (file in json_files) {
  # Read the JSON file
  
  json_data <- fromJSON(file)
  
  total_bases <- json_data$basepair_counts$output
  output_reads <- json_data$read_counts$output
  
  avg_length <- round(total_bases / (output_reads * 2),digits = 4)  # Paired-end
  sample <- extracted_pattern <- sub(".*24JX-([0-9-]+)_run1.*", "\\1", file)
  
  tmp <- c(sample, avg_length)
  avg_length_all <- rbind(avg_length_all,tmp)
  
}

rownames(avg_length_all) <- NULL
colnames(avg_length_all) <- c("sample","avg_length/read")
avg_length_all <- as.data.frame(avg_length_all )

DT::datatable(avg_length_all,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Average length after trimming'),options = list(pageLength = 10) )
```


# Counting sites

## cutoff: read depth 10, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=5}

coverage <- 10
mepct <- 10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/processed/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_cov",coverage,"_mepct",mepct),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",coverage,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

### site overlap
file_list <- list.files(path =  folder_countsites,pattern = paste0("temp_with_id_cov",coverage,"_mepct",mepct),full.names = T)

overlap_count <- count_overlap(file_list)

cat(sprintf("The number of overlapping site among 4 samples is %s\n",overlap_count))


```


## cutoff: read depth 10, methelytion level = 20%

```{r warning=F, message=F, fig.height=5, fig.width=5}

coverage <- 10
mepct <-20

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/processed/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_cov",coverage,"_mepct",mepct),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",coverage,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

### site overlap
file_list <- list.files(path =  folder_countsites,pattern = paste0("temp_with_id_cov",coverage,"_mepct",mepct),full.names = T)

overlap_count <- count_overlap(file_list)

cat(sprintf("The number of overlapping site among 4 samples is %s\n",overlap_count))


```


## cutoff: read depth 10, methelytion level = 30%


```{r warning=F, message=F, fig.height=5, fig.width=5}

coverage <- 10
mepct <-30

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/processed/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_cov",coverage,"_mepct",mepct),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",coverage,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

### site overlap
file_list <- list.files(path =  folder_countsites,pattern = paste0("temp_with_id_cov",coverage,"_mepct",mepct),full.names = T)

overlap_count <- count_overlap(file_list)

cat(sprintf("The number of overlapping site among 4 samples is %s\n",overlap_count))

```


## cutoff: read depth 20, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=5}

coverage <- 20
mepct <-10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/processed/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_cov",coverage,"_mepct",mepct),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",coverage,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

### site overlap
file_list <- list.files(path =  folder_countsites,pattern = paste0("temp_with_id_cov",coverage,"_mepct",mepct),full.names = T)

overlap_count <- count_overlap(file_list)

cat(sprintf("The number of overlapping site among 4 samples is %s\n",overlap_count))

```

## cutoff: read depth 30, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=5}

coverage <- 30
mepct <-10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/processed/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_cov",coverage,"_mepct",mepct),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",coverage,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

### site overlap
file_list <- list.files(path =  folder_countsites,pattern = paste0("temp_with_id_cov",coverage,"_mepct",mepct),full.names = T)

overlap_count <- count_overlap(file_list)

cat(sprintf("The number of overlapping site among 4 samples is %s\n",overlap_count))

```


## cutoff: read depth 40, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=5}

coverage <- 40
mepct <-10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace/count_sites/pileup/processed/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_cov",coverage,"_mepct",mepct),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",coverage,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

### site overlap
file_list <- list.files(path =  folder_countsites,pattern = paste0("temp_with_id_cov",coverage,"_mepct",mepct),full.names = T)

overlap_count <- count_overlap(file_list)

cat(sprintf("The number of overlapping site among 4 samples is %s\n",overlap_count))

```