---
title: "Quality Control -- data round 5, from Jiahao (v3 pipeline)"
author: "XSun"
date: "2026-08-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

Round 5 (INLINE library; JLK07=Mfs_NI/RIH014, JLK08=Mfs_LPS/RIH004,
JLK09=Mfs_LPS/RIH007) 

```{r setup, message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
library(knitr)
library(kableExtra)
library(DT)

BASE <- "/project/xinhe/xsun/camseq/processed_new/round5_JL_merged"
READS_DIRS <- c(
  "/project/xinhe/xsun/camseq/processed_new/round5_JL",         # JLK07
  "/project/xinhe/xsun/camseq/processed_new/round5_JL_JLK0809"  # JLK08 + JLK09
)
SAMPLES <- c("JLK07", "JLK08", "JLK09")
```

# Sample information

| Sample | Condition        | NEB index |
|--------|-------------------|-----------|
| JLK07  | Mfs_NI/RIH014     | idx25     |
| JLK08  | Mfs_LPS/RIH004    | idx26     |
| JLK09  | Mfs_LPS/RIH007    | idx27     |

# MultiQC

[MultiQC report -- round 5, v3 pipeline](https://xsun1229.github.io/camseq-m6a/multiqc_round5_JL_v3.html)

# Read processing summary

```{r message=FALSE, warning=FALSE}
DT::datatable(
  qc_read_counts(READS_DIRS),
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "Reads at each pipeline stage"
  ),
  options = list(pageLength = 10)
)
```

# Site calling QC

Swept over binomial p-value cutoff x methylation-ratio (`Uncon/Depth`) cutoff,
with the pipeline's other two defaults held fixed (`min_uncon=1`,
`min_depth=10`), focusing on per-sample results for now (no
pooled/replicates-merged view). Motif plots show the top 10 motifs by count
for both the 5-mer (`motif_flanking=2`, the pipeline's own annotation) and
the derived +-1nt 3-mer; colors are assigned once per motif string and
reused everywhere, so the same motif is always the same color across every
plot in this doc.

## Genome sites

Number of called sites across the full cutoff grid:

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=9}
qc_heatmap_plot(
  qc_sweep_counts(BASE, "genome", SAMPLES, c("0.05", "0.001"), c("0.05", "0.10", "0.20")),
  "Genome sites called, round 5"
)
```

### p < 0.05, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.05"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

for (pr in combn(SAMPLES, 2, simplify = FALSE)) {
  cs <- qc_common_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(cs, pr[1], pr[2], sprintf("common (intersection), p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
  us <- qc_union_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(us, pr[1], pr[2], sprintf("union, p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
}
```

### p < 0.05, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.05"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

for (pr in combn(SAMPLES, 2, simplify = FALSE)) {
  cs <- qc_common_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(cs, pr[1], pr[2], sprintf("common (intersection), p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
  us <- qc_union_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(us, pr[1], pr[2], sprintf("union, p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
}
```

### p < 0.05, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.05"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

for (pr in combn(SAMPLES, 2, simplify = FALSE)) {
  cs <- qc_common_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(cs, pr[1], pr[2], sprintf("common (intersection), p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
  us <- qc_union_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(us, pr[1], pr[2], sprintf("union, p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
}
```

### p < 0.001, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.001"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

for (pr in combn(SAMPLES, 2, simplify = FALSE)) {
  cs <- qc_common_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(cs, pr[1], pr[2], sprintf("common (intersection), p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
  us <- qc_union_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(us, pr[1], pr[2], sprintf("union, p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
}
```

### p < 0.001, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.001"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

for (pr in combn(SAMPLES, 2, simplify = FALSE)) {
  cs <- qc_common_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(cs, pr[1], pr[2], sprintf("common (intersection), p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
  us <- qc_union_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(us, pr[1], pr[2], sprintf("union, p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
}
```

### p < 0.001, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.001"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

for (pr in combn(SAMPLES, 2, simplify = FALSE)) {
  cs <- qc_common_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(cs, pr[1], pr[2], sprintf("common (intersection), p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
  us <- qc_union_sites(dt_persample, pr[1], pr[2], r, p)
  print(qc_methylation_scatter(us, pr[1], pr[2], sprintf("union, p<%s, methylation>=%s%%", p, as.numeric(r) * 100)))
}
```

## Genes (rRNA reference) sites

Secondary reference (rRNA depletion target) -- site counts only, across the
same cutoff grid:

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=9}
qc_heatmap_plot(
  qc_sweep_counts(BASE, "genes", SAMPLES, c("0.05", "0.001"), c("0.05", "0.10", "0.20")),
  "Genes (rRNA) sites called, round 5"
)
```

### Sites along the 45S precursor rRNA (18S / 5.8S / 28S)

Same idea as a plot used in a different project
(`~/xsun.xin3/pumseq/percl_pipeline/bin/R0_Sites_on_Ribosome_250125.R`,
pseudouridine/m1A sites along the 45S rRNA), adapted for our own m6A calls:
every covered position (depth>=10) plotted along the concatenated 45S
precursor, colored by whether that sample individually calls it significant
at the pipeline's default cutoffs (green) vs background (grey). Our "genes"
reference also has 17 copies of 5S rRNA (Pol III-transcribed, not part of
the 45S precursor), excluded here since there's no single linear coordinate
they share with 18S/5.8S/28S.

```{r warning=FALSE, message=FALSE, fig.height=8, fig.width=8}
rrna_dt <- fread("/project/xinhe/xsun/camseq/processed_new/round5_JL_merged/report_sites/rrna_45S_sites.tsv")
qc_rrna_lollipop_plot(rrna_dt, "Round 5: m6A sites on the 45S precursor rRNA")
```
