---
title: "Quality Control -- data round 3"
author: "XSun"
date: "2026-01-18"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r message=F, warning=F}

library(jsonlite)
library(dplyr)
library(gridExtra)
library(ggplot2)
library(grid) 
library(stringr)
library(tidyverse)
library(ggsci)

library(readr)
library(purrr)

process_motif_data <- function(file_list) {
  # Process files and calculate totals
  combined_data <- map_df(file_list, ~{
    message("\nProcessing: ", normalizePath(.x))
    
    # Extract sample name
    sample_name <- basename(.x) %>%
      str_extract("^sample\\d+")
    
    # Read and process data
    sample_data <- read_table(.x, 
                             col_names = c("motif", "count"), 
                             col_types = "cn",
                             skip_empty_rows = TRUE)
    
    total_reads <- sum(sample_data$count)
    
    processed <- sample_data %>%
      mutate(
        sample = sample_name,
        total_count = total_reads,
        percentage = (count / total_reads) * 100
      ) %>%
      select(sample, motif, percentage, total_count)
    
    return(processed)
  })
  
  # Create formatted labels and order data
  combined_data %>%
    mutate(
      sample_label = paste0(sample, " (n=", 
                           format(total_count, big.mark = ",", scientific = FALSE), ")"),
      sample_label = fct_reorder(sample_label, total_count)
    ) %>%
    arrange(desc(total_count)) %>%
    mutate(sample_label = factor(sample_label, levels = unique(sample_label)))
}

process_motif_data_merged <- function(file_list) {
  # Process files and calculate totals
  combined_data <- map_df(file_list, ~{
    message("\nProcessing: ", normalizePath(.x))
    
    # Extract sample name
    sample_name <- basename(.x) %>%
      stringr::str_extract("(?<=merged_)[^.]+")
    
    # Read and process data
    sample_data <- read_table(.x, 
                             col_names = c("motif", "count"), 
                             col_types = "cn",
                             skip_empty_rows = TRUE)
    
    total_reads <- sum(sample_data$count)
    
    processed <- sample_data %>%
      mutate(
        sample = sample_name,
        total_count = total_reads,
        percentage = (count / total_reads) * 100
      ) %>%
      select(sample, motif, percentage, total_count)
    
    return(processed)
  })
  
  # Create formatted labels and order data
  combined_data %>%
    mutate(
      sample_label = paste0(sample, " (n=", 
                           format(total_count, big.mark = ",", scientific = FALSE), ")"),
      sample_label = fct_reorder(sample_label, total_count)
    ) %>%
    arrange(desc(total_count)) %>%
    mutate(sample_label = factor(sample_label, levels = unique(sample_label)))
}

create_motif_plot <- function(processed_motif_data, 
                             title = "Motif Distribution with Total Read read_depth",
                             x_lab = "Percentage (%)",
                             y_lab = "Sample (Total site read_depth passing filter)",
                             point_size = 3,
                             text_size = 3) {
  ggplot(processed_motif_data, aes(x = percentage, y = sample_label)) +
    geom_point(
      aes(color = motif),
      size = point_size,
      position = position_jitterdodge(
        jitter.width = 0.2,
        jitter.height = 0.1,
        dodge.width = 0.7
      ),
      show.legend = FALSE
    ) +
    geom_text(
      aes(label = motif, color = motif),
      position = position_jitterdodge(
        jitter.width = 0.2,
        jitter.height = 0.1,
        dodge.width = 0.7
      ),
      size = text_size,
      vjust = -1.5,
      show.legend = FALSE
    ) +
   # scale_color_igv(guide = "none") +
    labs(
      title = title,
      x = x_lab,
      y = y_lab
    ) +
    theme_minimal() +
    theme(
      panel.grid.major.y = element_line(color = "grey90"),
      axis.text.y = element_text(size = 9, face = "bold"),
      axis.text.x = element_text(size = 10),
      plot.title = element_text(hjust = 0.5, size = 14)
    )
}



count_overlap <- function(file_list) {
  # Initialize an empty list to store the 'name' columns
  name_list <- list()
  
  # Read each file and store the 'name' column
  for(file in file_list){
    df_site <- read.table(file, header = T)
    df_site$name <- paste0(df_site$ref, "_", df_site$pos)
    name_list <- append(name_list, list(df_site$name))
  }
  
  # Find the overlap among all 'name' columns
  overlap <- Reduce(intersect, name_list)
  
  # Return the overlap count
  return(length(overlap))
}


```


# Sample information 

| Condition / Description | Sample ID            | i7 Index Name | i7 Index Seq | i5 Index Name | i5 Index Seq | i5 Rev Comp | Manufacturer | Input |
|------------------------|-----------------------|---------------|--------------|---------------|--------------|-------------|-----------|-------|
| HEK293T input          | JL-CAM-1(sample01)    | NEB S773      | ATCTGACC     | NEB S593      | CTCGACTT     | AAGTCGAG    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| Mfs_NI_RIH014 treat    | JL-CAM-2(sample02)    | NEB S714      | AACCAGAG     | NEB S520      | CTTCGGTT     | AACCGAAG    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| Mfs_LPS_RIH004 treat   | JL-CAM-3(sample03)    | NEB S761      | GTACCACA     | NEB S533      | CCACAACA     | TGTTGTGG    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| Mfs_LPS_RIH007 treat   | JL-CAM-4(sample04)    | NEB S721      | CGAGAGAA     | NEB S508      | GGAAGAGA     | TCTCTTCC    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| HEK293T treat          | JL-CAM-5(sample05)    | NEB S792      | CAGCATAC     | NEB S534      | GCGTTAGA     | TCTAACGC    | NEBNext® Multiplex Oligos for Illumina® | 10M |


Sample 2-4 are from the same cell type, so they should be similar. So we merged the M6A sites of these 3 samples later.


# Output from multiqc 

## Pipeline output 

https://xsun1229.github.io/camseq-m6a/multiqc_report_genome_all_round3.html


# Trimming

**Sample 01 is input, others are treated**

```{r}
json_folder <- "/project/xinhe/xsun/camseq/data/workspace_round3/report_reads/trimming/"

# List all JSON files in the folder
json_files <- list.files(json_folder, pattern = "*.json", full.names = TRUE)

avg_length_all <- c()
for (file in json_files) {
  # Read the JSON file
  
  json_data <- fromJSON(file)
  
  total_bases <- json_data$basepair_counts$output
  output_reads <- json_data$read_counts$output
  
  avg_length <- round(total_bases / (output_reads * 2),digits = 4)  # Paired-end
  sample <- sub(".*/(sample[0-9]+)_run1.*", "\\1", file)

  
  tmp <- c(sample, avg_length)
  avg_length_all <- rbind(avg_length_all,tmp)
  
}

rownames(avg_length_all) <- NULL
colnames(avg_length_all) <- c("sample","avg_length/read")
avg_length_all <- as.data.frame(avg_length_all )

DT::datatable(avg_length_all,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Average length after trimming'),options = list(pageLength = 10) )
```

The input sample (sample01) retains substantially longer reads after trimming, indicating minimal chemical damage.

Treated samples show shorter average read lengths, consistent with chemical treatment effects.


<!-- # Checking the read depth and methylation level for each sample -->


<!-- ```{r fig.height=9, fig.width=8} -->

<!-- dir_workspace <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/" -->


<!-- files <- list.files(dir_workspace, pattern = "\\.genome\\.tsv\\.gz$", full.names = TRUE) -->

<!-- p <- list() -->
<!-- for (file in files) { -->

<!--   base <- tools::file_path_sans_ext(basename(file)) -->
<!--   base <- sub(".genome.tsv", "", base) -->


<!--   # Step 1: Filter sites and add IDs, include only chr1 - chr22 -->
<!--   #dt <- fread(cmd = paste("cat", file), sep = "\t") -->
<!--   dt <- data.table::fread(file) -->
<!--   colnames(dt) <- c("ref", "pos", "strand", "convertedBaseCount", "unconvertedBaseCount") -->
<!--   dt$read_depth <- dt$convertedBaseCount + dt$unconvertedBaseCount -->
<!--   dt$methylation_level <- (dt$unconvertedBaseCount / dt$read_depth) * 100 -->

<!--   p[[base]] <- ggplot(dt, aes(x = read_depth, y = methylation_level)) + -->
<!--     geom_hex(bins = 50) + -->
<!--     theme_bw() + -->
<!--     labs( -->
<!--       title = "Read Depth vs Methylation Level", -->
<!--       x = "Read depth", -->
<!--       y = "Methylation level (%)" -->
<!--     ) + ggtitle(base) -->
<!-- } -->

<!-- all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]], nrow = 3) -->

<!-- ``` -->


# Motif counting

Motifs are counted within ±1 nucleotide of the candidate modified base. This corresponds to a 3-nt window centered on the site of interest

## Filter the M6A sites by site-level p-value

**Site-level p-value calculation**

For each site, the number of unconverted reads is modeled as a binomial random variable,
\( X \sim \mathrm{Binomial}(n, p_{\mathrm{bg}}) \),
where \( n \) is the total read depth at the site and \( p_{\mathrm{bg}} \) represents the background probability of observing an unconverted A due to random sequencing or reverse-transcription errors and incomplete chemical conversion. Sites are first filtered by a minimum read depth to ensure stable statistical estimation. 

The background probability is set to 0.005

<!-- Because the input/control sample contains too few high-coverage sites to reliably estimate background, \( p_{\mathrm{bg}} \) is estimated empirically from a large set of background sites in the treated sample that are unlikely to be modified (e.g., m⁶A-rare motifs or sites with low unconverted fractions). Specifically, \( p_{\mathrm{bg}} \) is calculated as \( \sum k / \sum n \), where \( k \) and \( n \) denote the unconverted read counts and total depth aggregated across background sites. For each tested site, a one-sided binomial p-value is computed as \( P(X \ge k) \), and sites passing both the depth filter and a stringent p-value cutoff are classified as significantly modified. -->


ref: PMID: 40247118 SI page3


### read depth 20, p-value < 0.0001 (The background probability is set to 0.005)


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  20
binomp <- "0.0001"

folder_countsites <- paste0("/project/xinhe/xsun/camseq/data/workspace_round3/count_sites_merged/processed_binomp_",binomp,"_1n/")

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_binomp",binomp),full.names = T)

processed_motif_data <- process_motif_data_merged(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", binomial p-value < ",binomp),
                  point_size = 4,
                  text_size = 3.5)

```


```{r warning=F, message=F, fig.height=4, fig.width=5}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_binomp",binomp),full.names = T)

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")


  sample <- basename(file_list[i]) %>%
    stringr::str_extract("(?<=merged_)[^.]+")
  

  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", binomial p-value < ",binomp), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  20
binomp <- "0.0001"

folder_countsites <- paste0("/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_binomp_",binomp,"_1n/")

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_binomp",binomp),full.names = T)

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", binomial p-value < ",binomp),
                  point_size = 4,
                  text_size = 3.5)

```


```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_binomp",binomp),full.names = T)

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")

  
  sample <- basename(file_list[i]) %>%
    str_extract("^sample\\d+")
  

  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", binomial p-value < ",binomp), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```



## Filter the M6A sites by read depth and methylation level


To reduce false positives and low-confidence sites, the following thresholds are applied before motif analysis:

| Filter                    | Threshold  |
| ------------------------- | ---------- |
| Minimum read depth        | ≥ 10/20/30/40 reads |
| Minimum methylation level | ≥ 10/20/30%      |


Definitions:

- Read depth: total number of reads covering a genomic position
- Methylation level:converted reads / total reads * 100%

Only sites passing both criteria are retained 


### cutoff: read depth 10, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  10
mepct <- 10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_1/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

# ### site overlap
# file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
# 
# overlap_count <- count_overlap(file_list)
# cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count))

```


```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")


  sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>%
      str_remove(".genome.tsv") %>%
      str_remove("^_+") %>%
      str_remove("_+$")


  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```


### cutoff: read depth 10, methelytion level = 20%

```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  10
mepct <-20

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_1/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

# ### site overlap
# file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
# 
# overlap_count <- count_overlap(file_list)
# cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count))



```

```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")


  sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>%
      str_remove(".genome.tsv") %>%
      str_remove("^_+") %>%
      str_remove("_+$")


  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

### cutoff: read depth 10, methelytion level = 30%


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  10
mepct <-30

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_1/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

# ### site overlap
# file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
# 
# overlap_count <- count_overlap(file_list)
# cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count))


```

```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")


  data <- data[!data$methylation < mepct,]

  sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>%
      str_remove(".genome.tsv") %>%
      str_remove("^_+") %>%
      str_remove("_+$")


  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

### cutoff: read depth 20, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  20
mepct <-10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_1/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

# ### site overlap
# file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
# 
# overlap_count <- count_overlap(file_list)
# cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count))


```

```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")


  data <- data[!data$methylation < mepct,]

  sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>%
      str_remove(".genome.tsv") %>%
      str_remove("^_+") %>%
      str_remove("_+$")


  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

### cutoff: read depth 30, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  30
mepct <-10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_1/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

# ### site overlap
# file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
# 
# overlap_count <- count_overlap(file_list)
# cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count))


```

```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")

  data <- data[!data$methylation < mepct,]

  sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>%
      str_remove(".genome.tsv") %>%
      str_remove("^_+") %>%
      str_remove("_+$")


  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

### cutoff: read depth 40, methelytion level = 10%


```{r warning=F, message=F, fig.height=5, fig.width=6}

read_depth <-  40
mepct <-10

folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_1/"

### motif percentage
file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

processed_motif_data <- process_motif_data(file_list)
create_motif_plot(processed_motif_data,
                  title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"),
                  point_size = 4,
                  text_size = 3.5)

# ### site overlap
# file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
# 
# overlap_count <- count_overlap(file_list)
# cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count))


```

```{r warning=F, message=F, fig.height=4, fig.width=16}

file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T)
file_list <- file_list[2:5]

p <- list()
for (i in 1:length(file_list)){

  # Read the data
  data <- read.table(file_list[i],
                   header = TRUE, sep = "\t")

  sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>%
      str_remove(".genome.tsv") %>%
      str_remove("^_+") %>%
      str_remove("_+$")


  p[[i]] <- ggplot(data, aes(x = methylation_mepct)) +
    geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
    labs(title = sample,
         x = "Methylation Percentage (%)",
         y = "Frequency") +
    theme_minimal()


}

combined <- arrangeGrob(
  grobs = p,
  nrow = 1,
  top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

<!-- ## +- 2 nucleosides -->

<!-- ### cutoff: read depth 10, methelytion level = 10% -->


<!-- ```{r warning=F, message=F, fig.height=5, fig.width=6} -->

<!-- read_depth <-  10 -->
<!-- mepct <- 10 -->

<!-- folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_2/" -->

<!-- ### motif percentage -->
<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- file_list <- file_list[2:5] -->

<!-- processed_motif_data <- process_motif_data(file_list) -->
<!-- create_motif_plot(processed_motif_data, -->
<!--                   title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"), -->
<!--                   point_size = 4, -->
<!--                   text_size = 3.5) -->

<!-- # ### site overlap -->
<!-- # file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- #  -->
<!-- # overlap_count <- count_overlap(file_list) -->
<!-- # cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count)) -->


<!-- ``` -->


<!-- ```{r warning=F, message=F, fig.height=4, fig.width=16} -->

<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->

<!-- p <- list() -->
<!-- for (i in 1:length(file_list)){ -->

<!--   # Read the data -->
<!--   data <- read.table(file_list[i], -->
<!--                    header = TRUE, sep = "\t") -->


<!--   sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>% -->
<!--       str_remove(".genome.tsv") %>% -->
<!--       str_remove("^_+") %>% -->
<!--       str_remove("_+$") -->


<!--   p[[i]] <- ggplot(data, aes(x = methylation_mepct)) + -->
<!--     geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + -->
<!--     labs(title = sample, -->
<!--          x = "Methylation Percentage (%)", -->
<!--          y = "Frequency") + -->
<!--     theme_minimal() -->


<!-- } -->

<!-- combined <- arrangeGrob( -->
<!--   grobs = p, -->
<!--   nrow = 1, -->
<!--   top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14)) -->
<!-- ) -->
<!-- grid.draw(combined) -->

<!-- ``` -->


<!-- ### cutoff: read depth 10, methelytion level = 20% -->

<!-- ```{r warning=F, message=F, fig.height=5, fig.width=6} -->

<!-- read_depth <-  10 -->
<!-- mepct <-20 -->

<!-- folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_2/" -->

<!-- ### motif percentage -->
<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- file_list <- file_list[2:5] -->

<!-- processed_motif_data <- process_motif_data(file_list) -->
<!-- create_motif_plot(processed_motif_data, -->
<!--                   title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"), -->
<!--                   point_size = 4, -->
<!--                   text_size = 3.5) -->

<!-- # ### site overlap -->
<!-- # file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- #  -->
<!-- # overlap_count <- count_overlap(file_list) -->
<!-- # cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count)) -->



<!-- ``` -->

<!-- ```{r warning=F, message=F, fig.height=4, fig.width=16} -->

<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->

<!-- p <- list() -->
<!-- for (i in 1:length(file_list)){ -->

<!--   # Read the data -->
<!--   data <- read.table(file_list[i], -->
<!--                    header = TRUE, sep = "\t") -->


<!--   sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>% -->
<!--       str_remove(".genome.tsv") %>% -->
<!--       str_remove("^_+") %>% -->
<!--       str_remove("_+$") -->


<!--   p[[i]] <- ggplot(data, aes(x = methylation_mepct)) + -->
<!--     geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + -->
<!--     labs(title = sample, -->
<!--          x = "Methylation Percentage (%)", -->
<!--          y = "Frequency") + -->
<!--     theme_minimal() -->


<!-- } -->

<!-- combined <- arrangeGrob( -->
<!--   grobs = p, -->
<!--   nrow = 1, -->
<!--   top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14)) -->
<!-- ) -->
<!-- grid.draw(combined) -->

<!-- ``` -->

<!-- ### cutoff: read depth 10, methelytion level = 30% -->


<!-- ```{r warning=F, message=F, fig.height=5, fig.width=6} -->

<!-- read_depth <-  10 -->
<!-- mepct <-30 -->

<!-- folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_2/" -->

<!-- ### motif percentage -->
<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- file_list <- file_list[2:5] -->

<!-- processed_motif_data <- process_motif_data(file_list) -->
<!-- create_motif_plot(processed_motif_data, -->
<!--                   title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"), -->
<!--                   point_size = 4, -->
<!--                   text_size = 3.5) -->

<!-- # ### site overlap -->
<!-- # file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- #  -->
<!-- # overlap_count <- count_overlap(file_list) -->
<!-- # cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count)) -->


<!-- ``` -->

<!-- ```{r warning=F, message=F, fig.height=4, fig.width=16} -->

<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->

<!-- p <- list() -->
<!-- for (i in 1:length(file_list)){ -->

<!--   # Read the data -->
<!--   data <- read.table(file_list[i], -->
<!--                    header = TRUE, sep = "\t") -->


<!--   sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>% -->
<!--       str_remove(".genome.tsv") %>% -->
<!--       str_remove("^_+") %>% -->
<!--       str_remove("_+$") -->


<!--   p[[i]] <- ggplot(data, aes(x = methylation_mepct)) + -->
<!--     geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + -->
<!--     labs(title = sample, -->
<!--          x = "Methylation Percentage (%)", -->
<!--          y = "Frequency") + -->
<!--     theme_minimal() -->


<!-- } -->

<!-- combined <- arrangeGrob( -->
<!--   grobs = p, -->
<!--   nrow = 1, -->
<!--   top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14)) -->
<!-- ) -->
<!-- grid.draw(combined) -->

<!-- ``` -->

<!-- ### cutoff: read depth 20, methelytion level = 10% -->


<!-- ```{r warning=F, message=F, fig.height=5, fig.width=6} -->

<!-- read_depth <-  20 -->
<!-- mepct <-10 -->

<!-- folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_2/" -->

<!-- ### motif percentage -->
<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- file_list <- file_list[2:5] -->

<!-- processed_motif_data <- process_motif_data(file_list) -->
<!-- create_motif_plot(processed_motif_data, -->
<!--                   title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"), -->
<!--                   point_size = 4, -->
<!--                   text_size = 3.5) -->

<!-- # ### site overlap -->
<!-- # file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- #  -->
<!-- # overlap_count <- count_overlap(file_list) -->
<!-- # cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count)) -->


<!-- ``` -->

<!-- ```{r warning=F, message=F, fig.height=4, fig.width=16} -->

<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->

<!-- p <- list() -->
<!-- for (i in 1:length(file_list)){ -->

<!--   # Read the data -->
<!--   data <- read.table(file_list[i], -->
<!--                    header = TRUE, sep = "\t") -->


<!--   sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>% -->
<!--       str_remove(".genome.tsv") %>% -->
<!--       str_remove("^_+") %>% -->
<!--       str_remove("_+$") -->


<!--   p[[i]] <- ggplot(data, aes(x = methylation_mepct)) + -->
<!--     geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + -->
<!--     labs(title = sample, -->
<!--          x = "Methylation Percentage (%)", -->
<!--          y = "Frequency") + -->
<!--     theme_minimal() -->


<!-- } -->

<!-- combined <- arrangeGrob( -->
<!--   grobs = p, -->
<!--   nrow = 1, -->
<!--   top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14)) -->
<!-- ) -->
<!-- grid.draw(combined) -->

<!-- ``` -->

<!-- ### cutoff: read depth 30, methelytion level = 10% -->


<!-- ```{r warning=F, message=F, fig.height=5, fig.width=6} -->

<!-- read_depth <-  30 -->
<!-- mepct <-10 -->

<!-- folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_2/" -->

<!-- ### motif percentage -->
<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- file_list <- file_list[2:5] -->

<!-- processed_motif_data <- process_motif_data(file_list) -->
<!-- create_motif_plot(processed_motif_data, -->
<!--                   title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"), -->
<!--                   point_size = 4, -->
<!--                   text_size = 3.5) -->

<!-- # ### site overlap -->
<!-- # file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- #  -->
<!-- # overlap_count <- count_overlap(file_list) -->
<!-- # cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count)) -->


<!-- ``` -->

<!-- ```{r warning=F, message=F, fig.height=4, fig.width=16} -->

<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->

<!-- p <- list() -->
<!-- for (i in 1:length(file_list)){ -->

<!--   # Read the data -->
<!--   data <- read.table(file_list[i], -->
<!--                    header = TRUE, sep = "\t") -->


<!--   sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>% -->
<!--       str_remove(".genome.tsv") %>% -->
<!--       str_remove("^_+") %>% -->
<!--       str_remove("_+$") -->


<!--   p[[i]] <- ggplot(data, aes(x = methylation_mepct)) + -->
<!--     geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + -->
<!--     labs(title = sample, -->
<!--          x = "Methylation Percentage (%)", -->
<!--          y = "Frequency") + -->
<!--     theme_minimal() -->


<!-- } -->

<!-- combined <- arrangeGrob( -->
<!--   grobs = p, -->
<!--   nrow = 1, -->
<!--   top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14)) -->
<!-- ) -->
<!-- grid.draw(combined) -->

<!-- ``` -->

<!-- ### cutoff: read depth 40, methelytion level = 10% -->


<!-- ```{r warning=F, message=F, fig.height=5, fig.width=6} -->

<!-- read_depth <-  40 -->
<!-- mepct <-10 -->

<!-- folder_countsites <- "/project/xinhe/xsun/camseq/data/workspace_round3/count_sites/pileup/processed_2/" -->

<!-- ### motif percentage -->
<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("motif_summary_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- file_list <- file_list[2:5] -->

<!-- processed_motif_data <- process_motif_data(file_list) -->
<!-- create_motif_plot(processed_motif_data, -->
<!--                   title = paste0("read depth = ",read_depth,", methylation level = ",mepct,"%"), -->
<!--                   point_size = 4, -->
<!--                   text_size = 3.5) -->

<!-- # ### site overlap -->
<!-- # file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->
<!-- #  -->
<!-- # overlap_count <- count_overlap(file_list) -->
<!-- # cat(sprintf("The number of overlapping site for the two sample is %s\n",overlap_count)) -->


<!-- ``` -->

<!-- ```{r warning=F, message=F, fig.height=4, fig.width=16} -->

<!-- file_list <- list.files(path =  folder_countsites,pattern = paste0("with_id_read",read_depth,"_mepct",mepct),full.names = T) -->

<!-- p <- list() -->
<!-- for (i in 1:length(file_list)){ -->

<!--   # Read the data -->
<!--   data <- read.table(file_list[i], -->
<!--                    header = TRUE, sep = "\t") -->


<!--   sample <- basename(file_list[i]) %>% str_remove("_with_id_read\\d+_mepct\\d+\\.tsv$") %>% -->
<!--       str_remove(".genome.tsv") %>% -->
<!--       str_remove("^_+") %>% -->
<!--       str_remove("_+$") -->


<!--   p[[i]] <- ggplot(data, aes(x = methylation_mepct)) + -->
<!--     geom_histogram(binwidth = 5, fill = "skyblue", color = "black") + -->
<!--     labs(title = sample, -->
<!--          x = "Methylation Percentage (%)", -->
<!--          y = "Frequency") + -->
<!--     theme_minimal() -->


<!-- } -->

<!-- combined <- arrangeGrob( -->
<!--   grobs = p, -->
<!--   nrow = 1, -->
<!--   top = textGrob(paste0("Cutoff: read depth =",read_depth,", methlation level = ",mepct), gp = gpar(fontsize = 14)) -->
<!-- ) -->
<!-- grid.draw(combined) -->

<!-- ``` -->


