---
title: "Quality Control -- data round 4, from Jiahao (v3 pipeline)"
author: "XSun"
date: "2026-08-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

Single sample (`JLE21`, INLINE library, Mfs_NI/RIH014). See
[QC -- round 3](QC_round3_JL_v3.html) for the full explanation of the
binomial p-value / methylation-ratio sweep used below -- with only one
sample, the "per sample" and "pooled" views are mathematically identical, so
this doc only shows one, and there's no cross-sample overlap to compute.

```{r setup, message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
library(knitr)
library(kableExtra)
library(DT)

BASE <- "/project/xinhe/xsun/camseq/processed_new/round4_JL"
SAMPLES <- c("JLE21")
```

# Sample information

`JLE21`, i7 index TCGAAGAT, INLINE library type, Mfs_NI/RIH014.

# MultiQC

[MultiQC report -- round 4, v3 pipeline](https://xsun1229.github.io/camseq-m6a/multiqc_round4_JL_v3.html)

# Read processing summary

```{r message=FALSE, warning=FALSE}
DT::datatable(
  qc_read_counts(BASE),
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "Reads at each pipeline stage"
  ),
  options = list(pageLength = 10)
)
```

# Site calling QC

## Genome sites

```{r warning=FALSE, message=FALSE}
grid <- expand.grid(p = c("0.05", "0.001"), r = c("0.05", "0.10", "0.20"), stringsAsFactors = FALSE)
rows <- lapply(seq_len(nrow(grid)), function(i) {
  p <- grid$p[i]; r <- grid$r[i]
  dt <- qc_read_sweep(BASE, "genome", "pooled", p, r)
  data.table(p = p, methylation_ratio = r, n_sites = nrow(dt))
})
kable(rbindlist(rows)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=7}
qc_motif_plot(qc_motif_table_pooled(qc_read_sweep(BASE, "genome", "pooled", "0.05", "0.05")),
              title = "JLE21, p<0.05, methylation>=5%")
```

### p < 0.001, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=7}
qc_motif_plot(qc_motif_table_pooled(qc_read_sweep(BASE, "genome", "pooled", "0.001", "0.20")),
              title = "JLE21, p<0.001, methylation>=20%")
```

## Genes (rRNA reference) sites

```{r warning=FALSE, message=FALSE}
rows <- lapply(seq_len(nrow(grid)), function(i) {
  p <- grid$p[i]; r <- grid$r[i]
  dt <- qc_read_sweep(BASE, "genes", "pooled", p, r)
  data.table(p = p, methylation_ratio = r, n_sites = nrow(dt))
})
kable(rbindlist(rows)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```
