---
title: "Quality Control -- data round 1+2, from Pingluan (v3 pipeline)"
author: "XSun"
date: "2026-08-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

Round 2 is deeper re-sequencing of the same libraries as round 1 (`12-1-2`,
`12-1-3`, `12-1-4`, `12-1-7`, `12-1-8`); the v3 config merges both rounds'
FASTQs per sample before mapping. See [Pipeline overview](pipeline_overview.html)
for what changed vs. the original `QC.Rmd`/`QC_2rounds.Rmd`/`QC_deeper.Rmd`,
and [QC -- round 3](QC_round3_JL_v3.html) for the full explanation of the
binomial p-value / methylation-ratio sweep used below.

```{r setup, message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
library(knitr)
library(kableExtra)
library(DT)

BASE <- "/project/xinhe/xsun/camseq/processed_new/round1_2_PW"
SAMPLES <- c("12-1-1", "12-1-2", "12-1-3", "12-1-4", "12-1-7", "12-1-8")
```

# Sample information

Only these 6 of the original 8 round-1 samples were carried into the v3 run
(`12-1-5`/`12-1-6` were dropped -- see [Project overview](pipeline_overview.html)):

| Sample Name | i7 Index Sequence   | i5 Index Sequence   | i5 reverse complement | Index Cat no | Description         |
|-------------|---------------------|----------------------|-----------------------|--------------|----------------------|
| 12-1-1      | ATTACTCG            | TATAGCCT             | AGGCTATA               | NEB#E7600    | 18-QTL treatment     |
| 12-1-2      | TCCGGAGA            | TATAGCCT             | AGGCTATA               | NEB#E7600    | 28-QTL treatment     |
| 12-1-3      | TTCCTCCT            | TTGCTTGC             | GCAAGCAA               | NEB#E7780    | 18-QTL untreatment   |
| 12-1-4      | TGCTTGCT            | TTGCTTGC             | GCAAGCAA               | NEB#E7780    | 28-QTL untreatment   |
| 12-1-7      | AAGGCGTA & AATCCAGC | TGTGGCAT & ACTACGGA  | ATGCCACA & TCCGTAGT     | NEB#E7780    | 18-QTL treatment     |
| 12-1-8      | CTGATGAG            | TGTGGCAT & ACTACGGA  | ATGCCACA & TCCGTAGT     | NEB#E7780    | 28-QTL treatment     |

# MultiQC

[MultiQC report -- round 1+2, v3 pipeline](https://xsun1229.github.io/camseq-m6a/multiqc_round1_2_PW_v3.html)

# Read processing summary

```{r message=FALSE, warning=FALSE}
DT::datatable(
  qc_read_counts(BASE),
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "Reads at each pipeline stage"
  ),
  options = list(pageLength = 10)
)
```

# Site calling QC

Swept over binomial p-value cutoff x methylation-ratio (`Uncon/Depth`) cutoff,
with the pipeline's other two defaults held fixed (`min_uncon=1`,
`min_depth=10`), focusing on per-sample results for now (no
pooled/replicates-merged view). Motif plots show the top 10 motifs by count
for both the 5-mer (`motif_flanking=2`, the pipeline's own annotation) and
the derived +-1nt 3-mer; colors are assigned once per motif string and
reused everywhere, so the same motif is always the same color across every
plot in this doc.

## Genome sites

Number of called sites across the full cutoff grid:

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=9}
qc_heatmap_plot(
  qc_sweep_counts(BASE, "genome", SAMPLES, c("0.05", "0.001"), c("0.05", "0.10", "0.20")),
  "Genome sites called, round 1+2"
)
```

### p < 0.05, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
p <- "0.05"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
p <- "0.05"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
p <- "0.05"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
p <- "0.001"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
p <- "0.001"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=6, fig.width=7}
p <- "0.001"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 5),
              title = paste0("5-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p), width = 3),
              title = paste0("3-mer, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Genes (rRNA reference) sites

Secondary reference (rRNA depletion target) -- site counts only, across the
same cutoff grid:

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=9}
qc_heatmap_plot(
  qc_sweep_counts(BASE, "genes", SAMPLES, c("0.05", "0.001"), c("0.05", "0.10", "0.20")),
  "Genes (rRNA) sites called, round 1+2"
)
```
