---
title: "Quality Control -- data round 4, from Jiahao (v3 pipeline)"
author: "XSun"
date: "2026-08-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

Single sample (`JLE21`, INLINE library, Mfs_NI/RIH014). See
[QC -- round 3](QC_round3_JL_v3.html) for the full explanation of the
binomial p-value / methylation-ratio sweep used below. With only one sample
there's no cross-sample overlap to compute, and (for now) these docs focus on
per-sample results rather than a pooled/replicates-merged view.

```{r setup, message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
library(knitr)
library(kableExtra)
library(DT)

BASE <- "/project/xinhe/xsun/camseq/processed_new/round4_JL"
SAMPLES <- c("JLE21")
```

# Sample information

`JLE21`, i7 index TCGAAGAT, INLINE library type, Mfs_NI/RIH014.

# MultiQC

[MultiQC report -- round 4, v3 pipeline](https://xsun1229.github.io/camseq-m6a/multiqc_round4_JL_v3.html)

# Read processing summary

```{r message=FALSE, warning=FALSE}
DT::datatable(
  qc_read_counts(BASE),
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "Reads at each pipeline stage"
  ),
  options = list(pageLength = 10)
)
```

# Site calling QC

Motif plots are restricted to the top 10 motifs by count (a 5-mer has up to
4^4=256 possible values with the center fixed at "A", which clutters the
plot); both the 5-mer (`motif_flanking=2`, the pipeline's own annotation) and
the derived +-1nt 3-mer are shown.

## Genome sites

Number of called sites across the full cutoff grid:

```{r warning=FALSE, message=FALSE, fig.height=2.5, fig.width=9}
sc <- qc_sweep_counts(BASE, "genome", SAMPLES, c("0.05", "0.001"), c("0.05", "0.10", "0.20"))
qc_heatmap_plot(sc, "Genome sites called -- JLE21")
```

### p < 0.05, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=7}
dt <- qc_read_sweep(BASE, "genome", "persample", "0.05", "0.05")
qc_motif_plot(qc_motif_table(dt, SAMPLES, 0.05, 0.05, width = 5), title = "JLE21, 5-mer, p<0.05, methylation>=5%")
qc_motif_plot(qc_motif_table(dt, SAMPLES, 0.05, 0.05, width = 3), title = "JLE21, 3-mer, p<0.05, methylation>=5%")
```

### p < 0.001, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=7}
dt <- qc_read_sweep(BASE, "genome", "persample", "0.001", "0.20")
qc_motif_plot(qc_motif_table(dt, SAMPLES, 0.20, 0.001, width = 5), title = "JLE21, 5-mer, p<0.001, methylation>=20%")
qc_motif_plot(qc_motif_table(dt, SAMPLES, 0.20, 0.001, width = 3), title = "JLE21, 3-mer, p<0.001, methylation>=20%")
```

## Genes (rRNA reference) sites

Secondary reference (rRNA depletion target) -- site counts only:

```{r warning=FALSE, message=FALSE, fig.height=2.5, fig.width=9}
sc_genes <- qc_sweep_counts(BASE, "genes", SAMPLES, c("0.05", "0.001"), c("0.05", "0.10", "0.20"))
qc_heatmap_plot(sc_genes, "Genes (rRNA) sites called -- JLE21")
```
