---
title: "Quality Control -- data round 5, from Jiahao (v3 pipeline)"
author: "XSun"
date: "2026-08-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

Round 5 (INLINE library; JLK07=Mfs_NI/RIH014, JLK08=Mfs_LPS/RIH004,
JLK09=Mfs_LPS/RIH007) had to be run as two separate pipeline jobs (JLK07
alone, then JLK08+JLK09 together) due to a Snakemake directory-lock conflict,
then merged back into one site table (`round5_JL_merged`) -- see
[Pipeline overview](pipeline_overview.html) for why. Read-level stats
(trimming/mapping/dedup) still live in the two original job directories, so
this doc's MultiQC and read-count sections read from both. See
[QC -- round 3](QC_round3_JL_v3.html) for the full explanation of the
binomial p-value / methylation-ratio sweep used below.

```{r setup, message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
library(knitr)
library(kableExtra)
library(DT)

BASE <- "/project/xinhe/xsun/camseq/processed_new/round5_JL_merged"
READS_DIRS <- c(
  "/project/xinhe/xsun/camseq/processed_new/round5_JL",         # JLK07
  "/project/xinhe/xsun/camseq/processed_new/round5_JL_JLK0809"  # JLK08 + JLK09
)
SAMPLES <- c("JLK07", "JLK08", "JLK09")
```

# Sample information

| Sample | Condition        | NEB index |
|--------|-------------------|-----------|
| JLK07  | Mfs_NI/RIH014     | idx25     |
| JLK08  | Mfs_LPS/RIH004    | idx26     |
| JLK09  | Mfs_LPS/RIH007    | idx27     |

# MultiQC

[MultiQC report -- round 5, v3 pipeline](https://xsun1229.github.io/camseq-m6a/multiqc_round5_JL_v3.html)

# Read processing summary

```{r message=FALSE, warning=FALSE}
DT::datatable(
  qc_read_counts(READS_DIRS),
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "Reads at each pipeline stage"
  ),
  options = list(pageLength = 10)
)
```

# Site calling QC

## Genome sites

### p < 0.05, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.05"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.05"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.05"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.001"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.001"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=4, fig.width=7}
p <- "0.001"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Genes (rRNA reference) sites

Secondary reference (rRNA depletion target) -- site counts only, across the
same cutoff grid:

```{r warning=FALSE, message=FALSE}
grid <- expand.grid(p = c("0.05", "0.001"), r = c("0.05", "0.10", "0.20"), stringsAsFactors = FALSE)
rows <- lapply(seq_len(nrow(grid)), function(i) {
  p <- grid$p[i]; r <- grid$r[i]
  dt_persample <- qc_read_sweep(BASE, "genes", "persample", p, r)
  dt_pooled <- qc_read_sweep(BASE, "genes", "pooled", p, r)
  n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
  data.table(p = p, methylation_ratio = r, t(as.list(setNames(n_sites, SAMPLES))), pooled = nrow(dt_pooled))
})
kable(rbindlist(rows)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```
