---
title: "Quality Control -- data round 3, from Jiahao (v3 pipeline)"
author: "XSun"
date: "2026-08-12"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

This redoes the round 3 QC using the v3 pipeline re-run (see
[Pipeline overview](pipeline_overview.html) for what changed vs. the original
`QC_round3.Rmd`). The main methodological difference: instead of a manual
depth x methylation-% grid with a fixed background rate, sites are now called
by the pipeline's own `filter_sites.py`, which estimates `Bg` (background
conversion rate) **per sample** from that sample's own genome-wide site
coverage, and tests each site with a one-sided binomial p-value against it
(formulas in [Pipeline overview](pipeline_overview.html#how-bg-background-and-the-binomial-p-value-are-actually-computed)).
Here that's swept over a small p-value x methylation-ratio grid, in both a
"per sample" view and a "pooled" (replicates summed before testing) view.

```{r setup, message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
library(knitr)
library(kableExtra)
library(DT)

BASE <- "/project/xinhe/xsun/camseq/processed_new/round3_JL"
SAMPLES <- c("JL-CAM-01", "JL-CAM-02", "JL-CAM-03", "JL-CAM-04", "JL-CAM-05")
```

# Sample information

| Condition / Description | Sample ID            | i7 Index Name | i7 Index Seq | i5 Index Name | i5 Index Seq | i5 Rev Comp | Manufacturer | Input |
|------------------------|-----------------------|---------------|--------------|---------------|--------------|-------------|-----------|-------|
| HEK293T input          | JL-CAM-1(sample01)    | NEB S773      | ATCTGACC     | NEB S593      | CTCGACTT     | AAGTCGAG    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| Mfs_NI_RIH014 treat    | JL-CAM-2(sample02)    | NEB S714      | AACCAGAG     | NEB S520      | CTTCGGTT     | AACCGAAG    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| Mfs_LPS_RIH004 treat   | JL-CAM-3(sample03)    | NEB S761      | GTACCACA     | NEB S533      | CCACAACA     | TGTTGTGG    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| Mfs_LPS_RIH007 treat   | JL-CAM-4(sample04)    | NEB S721      | CGAGAGAA     | NEB S508      | GGAAGAGA     | TCTCTTCC    | NEBNext® Multiplex Oligos for Illumina® | 10M |
| HEK293T treat          | JL-CAM-5(sample05)    | NEB S792      | CAGCATAC     | NEB S534      | GCGTTAGA     | TCTAACGC    | NEBNext® Multiplex Oligos for Illumina® | 10M |

# MultiQC

Ran directly against `report_reads/` (cutadapt trimming, HISAT-3N mapping, and
UMICollapse dedup reports are all natively recognized -- no manual parsing
needed for this section):

[MultiQC report -- round 3, v3 pipeline](https://xsun1229.github.io/camseq-m6a/multiqc_round3_JL_v3.html)

# Read processing summary

From the pipeline's own `report_reads/read_counts_summary.tsv`:

```{r message=FALSE, warning=FALSE}
DT::datatable(
  qc_read_counts(BASE),
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "Reads at each pipeline stage"
  ),
  options = list(pageLength = 10)
)
```

# Site calling QC

Swept over binomial p-value cutoff x methylation-ratio (`Uncon/Depth`) cutoff,
with the pipeline's other two defaults held fixed (`min_uncon=1`,
`min_depth=10`). For each combination: number of sites called per sample and
pooled (replicates summed before testing), the motif distribution of called
sites (5-mer, `motif_flanking=2`, same as the pipeline default), and the
overlap of called sites across samples.

## Genome sites

### p < 0.05, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
p <- "0.05"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
p <- "0.05"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.05, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
p <- "0.05"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 5%

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
p <- "0.001"; r <- "0.05"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 10%

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
p <- "0.001"; r <- "0.10"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

### p < 0.001, methylation ratio >= 20%

```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=7}
p <- "0.001"; r <- "0.20"
dt_persample <- qc_read_sweep(BASE, "genome", "persample", p, r)
dt_pooled <- qc_read_sweep(BASE, "genome", "pooled", p, r)

n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
kable(data.table(sample = c(SAMPLES, "pooled (replicates merged)"), n_sites = c(n_sites, nrow(dt_pooled)))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

qc_motif_plot(qc_motif_table(dt_persample, SAMPLES, as.numeric(r), as.numeric(p)),
              title = paste0("per sample, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))
qc_motif_plot(qc_motif_table_pooled(dt_pooled),
              title = paste0("pooled, p<", p, ", methylation>=", as.numeric(r) * 100, "%"))

ov <- qc_overlap(dt_persample, SAMPLES, as.numeric(r), as.numeric(p))
cat(sprintf("All-way overlap (site called in every sample): %d\n", ov$all_way))
kable(ov$pairwise, caption = "Pairwise overlap (row/col: n sites shared)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Genes (rRNA reference) sites

Secondary reference (rRNA depletion target) -- site counts only, across the
same cutoff grid:

```{r warning=FALSE, message=FALSE}
grid <- expand.grid(p = c("0.05", "0.001"), r = c("0.05", "0.10", "0.20"), stringsAsFactors = FALSE)
rows <- lapply(seq_len(nrow(grid)), function(i) {
  p <- grid$p[i]; r <- grid$r[i]
  dt_persample <- qc_read_sweep(BASE, "genes", "persample", p, r)
  dt_pooled <- qc_read_sweep(BASE, "genes", "pooled", p, r)
  n_sites <- sapply(SAMPLES, function(s) nrow(qc_sample_passing(dt_persample, s, as.numeric(r), as.numeric(p))))
  data.table(p = p, methylation_ratio = r, t(as.list(setNames(n_sites, SAMPLES))), pooled = nrow(dt_pooled))
})
kable(rbindlist(rows)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```
