---
title: "Round 5: m6A sites in miRNA loci"
author: "XSun"
date: "2026-08-19"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: console
---

# What this is

Round 5's called m6A sites ([QC -- round 5](QC_round5_JL_v3.html)), intersected
with miRNA loci, per sample (JLK07, JLK08, JLK09) -- no merging across
samples.

**A note on terminology**, since this tripped us up while building this:
there are three distinct levels of "miRNA" region, and only two of them are
available in any standard annotation:

| Level | What it is | Available here? |
|---|---|---|
| pri-miRNA | The full-length primary Pol II transcript (often kb-scale, capped/polyadenylated) | **No** -- not systematically annotated anywhere (Ensembl, UCSC, miRBase all lack it); boundaries aren't well characterized for most loci |
| pre-miRNA (hairpin) | The ~60-100bp stem-loop released by Drosha cleavage from the pri-miRNA | **Yes** -- `premirna_hairpin` below |
| Mature miRNA (5p/3p arm) | The ~20-24nt product after Dicer processing | **Yes** -- `mature_mirna` below |

miRBase's own GFF3 calls the hairpin feature `miRNA_primary_transcript`,
which is misleading -- despite the name, it is **not** the true pri-miRNA;
miRBase's own file header says so explicitly ("these sequences do not
represent the full primary transcript, rather a predicted stem-loop portion
that includes the precursor miRNA").

**Source**: miRBase v23, GRCh38 primary assembly (`hsa.gff3`,
`NCBI_Assembly:GCA_000001405.15`) -- 1,858 pre-miRNA hairpin loci, 3,008
mature 5p/3p arm loci.

**Sites**: round 5's canonical filtered output
(`report_sites/filtered/genome.tsv.gz`, all 3 samples, pipeline default
cutoffs `min_uncon=1`, `min_depth=10`, `min_ratio=0.05`, `min_pval=0.05`),
intersected via `GenomicRanges::findOverlaps` (strand-aware) against each
resolution's loci separately.

Extraction script:
`/project/xinhe/xsun/camseq/2.qc/extract_mirna_sites.R` (in the pipeline
directory, not this site's repo).

```{r setup, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(DT)

MIRNA_DIR <- "/project/xinhe/xsun/camseq/processed_new/round5_JL_merged/report_sites/mirna"
SAMPLES <- c("JLK07", "JLK08", "JLK09")
RESOLUTIONS <- c(premirna_hairpin = "Pre-miRNA hairpin", mature_mirna = "Mature miRNA (5p/3p)")

read_one <- function(sample, resolution) {
  f <- file.path(MIRNA_DIR, sprintf("%s_%s_sites.tsv", sample, resolution))
  dt <- fread(f)
  dt[, `:=`(sample = sample, resolution = RESOLUTIONS[[resolution]])]
  dt
}

all_sites <- rbindlist(lapply(SAMPLES, function(s) {
  rbindlist(lapply(names(RESOLUTIONS), function(r) read_one(s, r)))
}))
```

# Summary

```{r message=FALSE, warning=FALSE}
summary_dt <- all_sites[, .(n_sites = .N, n_loci = uniqueN(mirna_name)), by = .(sample, resolution)]
kable(summary_dt) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=7}
ggplot(summary_dt, aes(x = resolution, y = n_sites, fill = sample)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = n_sites), position = position_dodge(width = 0.7), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = c(JLK07 = "#2a78d6", JLK08 = "#eb6834", JLK09 = "#1baf7a")) +
  labs(x = NULL, y = "m6A sites called", fill = NULL, title = "m6A sites in miRNA loci, round 5") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_classic(base_size = 13)
```

# Stricter cutoff: p<0.001, methylation ratio >=10%

Same intersection, but on a more stringent cutoff than the pipeline default
above (`min_uncon=1`, `min_depth=10` held fixed, `min_pval<0.001`,
`min_ratio>=0.10`), reading from the pre-computed
[QC sweep](QC_round5_JL_v3.html) output instead of the canonical filtered
file. JLK09's excess over JLK07/JLK08 persists at this stricter cutoff --
not just an artifact of the lenient default threshold picking up more
marginal/noisy calls.

```{r message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/qc_v3_helpers.R")
source("/project/xinhe/xsun/website/camseq-m6a/code/compare_db_helpers.R")

BASE <- "/project/xinhe/xsun/camseq/processed_new/round5_JL_merged"
P_STRICT <- "0.001"
R_STRICT <- "0.10"

mirna <- load_mirna_loci("/project/xinhe/xsun/camseq/data/reference/mirna/hsa.gff3")
dt_strict <- qc_read_sweep(BASE, "genome", "persample", P_STRICT, R_STRICT)
gr_strict <- GRanges(seqnames = dt_strict$chrom, ranges = IRanges(dt_strict$pos, dt_strict$pos), strand = dt_strict$strand)

strict_sites <- rbindlist(lapply(names(mirna), function(res_key) {
  gr_ref <- mirna[[res_key]]
  hits <- findOverlaps(gr_strict, gr_ref, ignore.strand = FALSE)
  overlap_dt <- cbind(dt_strict[queryHits(hits)], mirna_name = mcols(gr_ref)$name[subjectHits(hits)])
  rbindlist(lapply(SAMPLES, function(s) {
    passing <- qc_sample_passing(overlap_dt, s, as.numeric(R_STRICT), as.numeric(P_STRICT))
    passing <- unique(passing[, .(
      chrom, pos, strand, motif, mirna_name,
      Depth = get(paste0("Depth_", s)), Uncon = get(paste0("Uncon_", s)),
      methylation_pct = get(paste0("Ratio_", s)) * 100, pval = get(paste0("pval_", s))
    )])
    if (nrow(passing) == 0) return(NULL)
    passing[, `:=`(sample = s, resolution = if (res_key == "precursor") "Pre-miRNA hairpin" else "Mature miRNA (5p/3p)")]
    passing
  }))
}))

strict_summary <- strict_sites[, .(n_sites = .N, n_loci = uniqueN(mirna_name)), by = .(sample, resolution)]
kable(strict_summary) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=7}
ggplot(strict_summary, aes(x = resolution, y = n_sites, fill = sample)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = n_sites), position = position_dodge(width = 0.7), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = c(JLK07 = "#2a78d6", JLK08 = "#eb6834", JLK09 = "#1baf7a")) +
  labs(x = NULL, y = "m6A sites called", fill = NULL, title = "m6A sites in miRNA loci, round 5 (p<0.001, methylation>=10%)") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_classic(base_size = 13)
```

```{r message=FALSE, warning=FALSE}
DT::datatable(
  strict_sites[, .(sample, resolution, mirna_name, chrom, pos, strand, motif, Depth, Uncon, methylation_pct, pval)],
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "m6A sites in miRNA loci, round 5, p<0.001 & methylation>=10%"
  ),
  options = list(pageLength = 15),
  filter = "top"
) %>%
  DT::formatRound(columns = "methylation_pct", digits = 1) %>%
  DT::formatSignif(columns = "pval", digits = 3)
```

# Sites shared across samples

Same site (chrom, pos, strand) called in more than one sample -- worth a
second look, since agreement across independent samples is stronger
evidence than any single call.

```{r message=FALSE, warning=FALSE}
shared <- all_sites[, .(n_samples = uniqueN(sample), samples = paste(unique(sample), collapse = ",")),
                     by = .(chrom, pos, strand, motif, mirna_name, resolution)][n_samples > 1]
kable(shared[order(-n_samples)]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

# Published reference: HSPC d9 sites in miRNA loci

For context (not our data): the same miRNA-loci intersection applied to
[m6A RNA modifications are measured at single-base resolution across the mammalian transcriptome, PMID: 35288668](https://pubmed.ncbi.nlm.nih.gov/35288668/)'s HSPC d9 dataset (PMID 35288668,
single-nucleotide resolution, 23,833 total sites) -- a rough check of how
often *any* published m6A site falls in a miRNA locus at all.

```{r message=FALSE, warning=FALSE}
source("/project/xinhe/xsun/website/camseq-m6a/code/compare_db_helpers.R")

paper1_d9 <- load_paper1_sheet("/project/xinhe/xsun/camseq/pipeline/data/m6a_PMID35288668.xlsx", 8)
mirna <- load_mirna_loci("/project/xinhe/xsun/camseq/data/reference/mirna/hsa.gff3")

p1_chrom <- sub("-.*", "", paper1_d9$site)
p1_pos <- as.integer(sub(".*-", "", paper1_d9$site))
gr_p1 <- GRanges(seqnames = p1_chrom, ranges = IRanges(p1_pos, p1_pos))

hp_hits <- suppressWarnings(findOverlaps(gr_p1, mirna$precursor, ignore.strand = TRUE))
mt_hits <- suppressWarnings(findOverlaps(gr_p1, mirna$mature, ignore.strand = TRUE))

paper1_mirna <- rbind(
  cbind(paper1_d9[queryHits(hp_hits)], mirna_name = mcols(mirna$precursor)$name[subjectHits(hp_hits)], resolution = "Pre-miRNA hairpin"),
  cbind(paper1_d9[queryHits(mt_hits)], mirna_name = mcols(mirna$mature)$name[subjectHits(mt_hits)], resolution = "Mature miRNA (5p/3p)")
)
setnames(paper1_mirna, "m6a_pct", "methylation_pct")

kable(paper1_mirna, caption = "Paper 1 HSPC d9 sites overlapping a miRNA locus") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

# cross-check against round5's own cross-sample-shared miRNA loci above
p1_loci <- sub("-5p$|-3p$", "", paper1_mirna$mirna_name)
p1_loci <- unique(sub("^hsa-miR-", "hsa-mir-", p1_loci))
shared_loci <- unique(sub("-5p$|-3p$", "", shared$mirna_name))
shared_loci <- unique(sub("^hsa-miR-", "hsa-mir-", shared_loci))
in_both <- intersect(p1_loci, shared_loci)
if (length(in_both) > 0) {
  cat(sprintf(
    "%d of Paper 1's miRNA-overlapping loci (%s) also appear in round 5's own cross-sample-shared list above.\n",
    length(in_both), paste(in_both, collapse = ", ")
  ))
}
```

# All sites

Searchable/sortable -- filter by sample, resolution, or miRNA name.

```{r message=FALSE, warning=FALSE}
DT::datatable(
  all_sites[, .(sample, resolution, mirna_name, chrom, pos, strand, motif, Depth, Uncon, methylation_pct, pval)],
  caption = htmltools::tags$caption(
    style = "caption-side: left; text-align: left; color:black; font-size:150%;",
    "m6A sites in miRNA loci, round 5"
  ),
  options = list(pageLength = 15),
  filter = "top"
) %>%
  DT::formatRound(columns = "methylation_pct", digits = 1) %>%
  DT::formatSignif(columns = "pval", digits = 3)
```
