---
title: "trimming"
author: "XSun"
date: "2025-02-26"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r message=F, warning=F}

library(jsonlite)
library(dplyr)
library(gridExtra)
library(ggplot2)
library(grid) 

```


# Commands

```{shell}

# Define the JSON report file
JSON_REPORT="report_reads/trimming/241217AVITI-He-lab-24JX-12-1-4.json"
echo "JSON report will be saved to: $JSON_REPORT"

# Define input and output files
INPUT_R1="raw_reads/241217AVITI-He-lab-24JX-12-1-4_R1.fastq.gz"
INPUT_R2="raw_reads/241217AVITI-He-lab-24JX-12-1-4_R2.fastq.gz"
OUTPUT_R1="intermediate/trimming/trimmed_reads/PE/241217AVITI-He-lab-24JX-12-1-4_R1.fq.gz"
OUTPUT_R2="intermediate/trimming/trimmed_reads/PE/241217AVITI-He-lab-24JX-12-1-4_R2.fq.gz"
DISCARDED_R1="intermediate/trimming/discarded_reads/241217AVITI-He-lab-24JX-12-1-4_R1.fq.gz"
DISCARDED_R2="intermediate/trimming/discarded_reads/241217AVITI-He-lab-24JX-12-1-4_R2.fq.gz"

    cutadapt \
        -a AGATCGGAAGAGCACACGTCTGAACTCCAGTCA \
        -A AGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGT \
        -j 8 \
        -m 20 \
        -o "$OUTPUT_R1" \
        -p "$OUTPUT_R2" \
        --too-short-output "$DISCARDED_R1" \
        --too-short-paired-output "$DISCARDED_R2" \
        --json="$JSON_REPORT" \
        "$INPUT_R1" "$INPUT_R2"

```


# Results

## Summary

```{r}
json_folder <- "/project/xinhe/xsun/camseq/report_reads/trimming"

# List all JSON files in the folder
json_files <- list.files(json_folder, pattern = "*.json", full.names = TRUE)

# Initialize an empty list to store results
results <- list()

# Loop through each JSON file
for (file in json_files) {
  # Read the JSON file
  json_data <- fromJSON(file)
  
  # Extract general statistics
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  total_reads <- json_data$read_counts$input
  reads_too_short <- json_data$read_counts$filtered$too_short
  reads_written <- json_data$read_counts$output
  reads_with_adapters_R1 <- json_data$read_counts$read1_with_adapter
  reads_with_adapters_R2 <- json_data$read_counts$read2_with_adapter
  total_basepairs_input <- json_data$basepair_counts$input
  total_basepairs_output <- json_data$basepair_counts$output
  
  # Store the results in a data frame
  results[[sample_name]] <- data.frame(
    Sample = sample_name,
    Total_Reads = total_reads,
    Reads_Too_Short = reads_too_short,
    Reads_Written = reads_written,
    Reads_With_Adapters_R1 = reads_with_adapters_R1,
    Reads_With_Adapters_R2 = reads_with_adapters_R2,
    Total_Basepairs_Input = total_basepairs_input,
    Total_Basepairs_Output = total_basepairs_output
  )
}

# Combine results into a single data frame
results_df <- do.call(rbind, results)

rownames(results_df) <- seq(1:nrow(results_df))

DT::datatable(results_df,caption = htmltools::tags$caption(style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Trimming summary'),options = list(pageLength = 10) )
```

## Read Counts Summary


```{r fig.height=8, fig.width=16}

p <- list()
for (file in json_files){
  
  json_data <- fromJSON(file)
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  
  read_counts <- data.frame(
    Category = c("Input", "Output", "Too Short"),
    Count = c(
      json_data$read_counts$input,
      json_data$read_counts$output,
      json_data$read_counts$filtered$too_short
    )
  )
  
  # Bar plot
  p[[length(p)+1]] <- ggplot(read_counts, aes(x = Category, y = Count)) +
    geom_col(fill = "skyblue") +
    geom_text(aes(label = scales::comma(Count)), vjust = -0.5) +
    labs(title = sample_name , y = "Count") +
    scale_y_continuous(labels = scales::comma)
  
  
}

combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("Read Counts Before/After Trimming", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

## Adapter Detection

```{r fig.height=8, fig.width=16}

p <- list()
for (file in json_files){
  
  json_data <- fromJSON(file)
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  
  adapter_counts <- data.frame(
    Read = c("Read 1", "Read 2"),
    Count = c(
      json_data$read_counts$read1_with_adapter,
      json_data$read_counts$read2_with_adapter
    )
  )
  
  # Bar plot
  p[[length(p)+1]] <- ggplot(adapter_counts, aes(x = Read, y = Count)) +
    geom_col(fill = "salmon") +
    geom_text(aes(label = scales::comma(Count)), vjust = -0.5) +
    labs(title = sample_name , y = "Count") 
    #labs(title = "Reads Containing Adapters", y = "Count")
  
  
}


combined <- arrangeGrob(
  grobs = p,
  nrow = 2,
  top = textGrob("Reads Containing Adapters", gp = gpar(fontsize = 14))
)
grid.draw(combined)

```

## Trimmed Length Distribution

The expected count is a theoretical prediction of random adapter matches:

$$
\text{Expected Count} = \text{Total Reads} \times P(\text{Random Match})
$$

Where:
- \(\text{Total Reads}\) is the total number of reads in the dataset.
- \(P(\text{Random Match})\) is the probability of a random sequence matching the adapter.

- If the blue line is much higher than the red line at certain lengths, it indicates true adapter trimming at those lengths.
- If the blue and red lines overlap, it suggests that the trimming at those lengths is likely due to random chance, not adapter presence.

```{r fig.height=8, fig.width=16}

p1 <- list()
p2 <- list()
for (file in json_files){
  
  json_data <- fromJSON(file)
  
  sample_name <- tools::file_path_sans_ext(basename(file))  # Get sample name from file name
  sample_name <- gsub(pattern = "241217AVITI-He-lab-24JX-",replacement = "",x =sample_name)
  
  process_trimmed <- function(adapter_data) {
    df <- adapter_data$three_prime_end$trimmed_lengths[[1]]
    df$total_observed <- sapply(df$counts, function(x) sum(unlist(x)))
    df$len <- as.numeric(df$len)
    df
  }
  
  # Process Read 1 and Read 2
  trimmed_read1 <- process_trimmed(json_data$adapters_read1)
  trimmed_read2 <- process_trimmed(json_data$adapters_read2)
  
  # Plot for Read 1
  p1[[length(p1)+1]] <- ggplot(trimmed_read1, aes(x = len)) +
    geom_line(aes(y = total_observed, color = "Observed"), linewidth = 1) +
    geom_line(aes(y = expect, color = "Expected"), linewidth = 1, linetype = "dashed") +
    #labs(title = "Trimmed Length Distribution (Read 1)", x = "Length", y = "Count") +
    labs(title = sample_name, x = "Length", y = "Count") +
    scale_color_manual(values = c("Observed" = "blue", "Expected" = "red")) +
    theme_minimal()
  
  p2[[length(p2)+1]] <- ggplot(trimmed_read2, aes(x = len)) +
    geom_line(aes(y = total_observed, color = "Observed"), linewidth = 1) +
    geom_line(aes(y = expect, color = "Expected"), linewidth = 1, linetype = "dashed") +
    #labs(title = "Trimmed Length Distribution (Read 2)", x = "Length", y = "Count") +
    labs(title = sample_name, x = "Length", y = "Count") +
    scale_color_manual(values = c("Observed" = "blue", "Expected" = "red")) +
    theme_minimal()

}

combined1 <- arrangeGrob(
  grobs = p1,
  nrow = 2,
  top = textGrob("Trimmed Length Distribution (Read 1)", gp = gpar(fontsize = 14))
)
grid.draw(combined1)
```

```{r fig.height=8, fig.width=16}
combined2 <- arrangeGrob(
  grobs = p2,
  nrow = 2,
  top = textGrob("Trimmed Length Distribution (Read 2)", gp = gpar(fontsize = 14))
)
grid.draw(combined2)

```